---
geometry: margin=1in
month: "`r format(Sys.Date(), '%B')`"
year: "`r format(Sys.Date(), '%Y')`"
preamble: |
output:
  sa4ss::techreport_pdf:
    default
  bookdown::pdf_document2:
    keep_tex: true
lang: en
papersize: letter
---

```{r opts, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::knit_hooks$set(plot = function(x,options) {
      base = knitr::opts_knit$get('base.url')
      if (is.null(base)) base = ''
      alt = ifelse (is.null(options$alt),"",options$alt)
      cap = ifelse (is.null(options$caption),"",options$caption)
      if (alt != ""){
        sprintf('![%s](%s%s "%s")', cap, base, x, alt)
      } else {
        sprintf('![%s](%s%s)', cap, base, x)
        }
  })

load("00opts.Rdata")

if(file.exists("00mod.Rdata")){
  load("00mod.Rdata")
} else {
  print("Model output not being read. Please run the read_model() function.")
}

spp = 'Black Rockfish'
Spp = 'Black Rockfish'

if(file.exists("00mod.Rdata")){
  load("00mod.Rdata")
} else {
  print("Model output not being read. Please run the read_model() function.")
}

base_name = "Reference model"
 mod_loc = file.path(
   "C:/Users/Jason.Cope/Documents/Github/Sebastes_melanops_WA/Document/models",
   base_name)
   
#base_name = "Reference model"
#mod_loc = file.path(
#   "C:/Users/copej/Documents/GitHub/Sebastes_melanops_WA/Document/models",
#   base_name)

bridge_fig_loc = "C:/Users/Jason.Cope/Documents/Github/Sebastes_melanops_WA/Document/figures/Bridge"

#stock_fig_loc = "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/figures/stock"

#bio_fig_loc = "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/figures/biology_plots"

#index_fig_loc = "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/figures/index"

#sensi_data_fig_loc = "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/figures/sensi_data"

#sensi_modspex_fig_loc = "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/figures/sensi_modspex"

#likeprof_fig_loc = "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/figures/likelihoods"

#retro_loc = "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/figures/retro"

source("derived_quants.R")

```

<!--chapter:end:00a.Rmd-->

---
author:
  - name: Jason M. Cope
    code: 1
    first: J
    middle: M
    family: Cope
  - name: Tien-Shui Tsou
    code: 2
    first: T
    middle: '-'
    family: Tsou
  - name: Lisa K. Hillier
    code: 2
    first: L
    middle: K
    family: Hillier
  - name: Kristen M. Hinton
    code: 2
    first: K
    middle: M
    family: Hinton
  - name: Corey B. Niles
    code: 2
    first: C
    middle: B
    family: Niles
author_list: Cope, J.M., T.-. Tsou, L.K. Hillier, K.M. Hinton, C.B. Niles
affiliation:
  - code: 1
    address: Northwest Fisheries Science Center, U.S. Department of Commerce, National
      Oceanic and Atmospheric Administration, National Marine Fisheries Service, 2725
      Montlake Boulevard East, Seattle, Washington 98112
  - code: 2
    address: Washington Department of Fish and Wildlife, 600 Capital Way North, Olympia,
      Washington 98501
address:
  - ^1^Northwest Fisheries Science Center, U.S. Department of Commerce, National Oceanic
    and Atmospheric Administration, National Marine Fisheries Service, 2725 Montlake
    Boulevard East, Seattle, Washington 98112
  - ^2^Washington Department of Fish and Wildlife, 600 Capital Way North, Olympia,
    Washington 98501
---

<!--chapter:end:00authors.Rmd-->

---
bibliography:
  - jmcbf2023.bib
---

<!--chapter:end:00bibliography.Rmd-->

---
title: Status of Black Rockfish (_Sebastes melanops_) off Washington and federal waters in 2023
---

<!--chapter:end:00title.Rmd-->

\vspace{500cm}

<!--# Disclaimer{-}

_**These materials do not constitute a formal publication and are for information only. They are in a pre-review, pre-decisional state and should not be formally cited or reproduced. They are to be considered provisional and do not represent any determination or policy of NOAA or the Department of Commerce.**_-->

\pagebreak
\pagenumbering{roman}
\setcounter{page}{1}

\renewcommand{\thetable}{\roman{table}}
\renewcommand{\thefigure}{\roman{figure}}


\setlength\parskip{0.5em plus 0.1em minus 0.2em}

<!--chapter:end:01a.Rmd-->

```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through xxxx.")
```

# Executive summary{-}

## Stock{-}

The assessments described in this document apply to the black rockfish (*Sebastes melanops*) stocks that reside in the waters from Point Conception (34\textdegree27' N latitude) in the south to the U.S. boundary with Canada (approximately 48\textdegree30' N latitude) through the year `r endyr`.  Following the consensus recommendations from a preliminary stock assessment workshop in April 2015 (PFMC 2015), the stock assessment team (STAT) decided to prepare separate geographic stock assessments that are spatially stratified with boundaries at the CA/OR border (42\textdegree00' N latitude) and OR/WA border (46\textdegree16' N latitude).  

Black rockfish are also caught from the waters off British Columbia and Alaska, but there have not been any formal assessments of stock status for those areas.

This assessment reports the status of `r spp` (_`r spp.sci`_) off Oregon state using data through `r endyr`. `r Spp` are also found in California (their core range) and Washington waters of the U.S. West Coast, and those are treated in separate stock assessments given different mangement considerations and exploitation histories. There is substantial biogeographic separation in the populations off Oregon and Washington, thus justifying separation of those populations into different management units and stock assessments.

## Removals{-}

Black rockfish are caught by a wide variety of gear types and in recent decades have been a very important target species for recreational charter-boats and private sport anglers in Washington and Oregon, and to a lesser extent in California.  In recent years the recreational fishery has accounted for most of the black rockfish catches (Figure ES-1 to Figure ES-3).  Black rockfish can also be an important component of nearshore commercial fisheries, either as incidental catch by the troll fishery for salmon or as directed catch by jig fisheries for groundfish.  Further, in California and Oregon there are nearshore fisheries that catch and sell fish live for the restaurant trade.  Washington closed nearshore commercial fisheries in state water in late 1990s and never allowed the live-fish fishery to develop. In all states there have been almost no trawl-caught landings of black rockfish in recent years (Table ES-1), but trawl landings in the past were substantial (Figure ES-1 to Figure ES-3).

Detailed reports of commercial landings of black rockfish are generally unavailable prior to 1981, when the Pacific Fishery Information Network (PacFIN) database began.  The catch series prior to 1981 for these assessments were derived by applying available estimates or assumed values for the proportion of black rockfish landings in reported landings of rockfish.  Observer data, which are available only for the past decade, indicate low levels of discarding of black rockfish, generally less than 2% of total catch.

Because of their nearshore distribution and low abundance compared to other rockfish species, black rockfish are unlikely to have ever comprised a large percentage of rockfish landings, but it seems quite certain that they have been more than a trivial component for many years.  Black rockfish were one of only four rockfish species mentioned by scientific name in reports of rockfish landings in Oregon during the 1940s, and they were one of only six rockfish species mentioned by scientific name in reports of rockfish landings in California during the same period. Mentions of black rockfish extend back before the year 1900 in Washington.


\clearpage

\input{tex_tables/a_Catches_ES.tex}


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the reference model where catches in metric tons by fleet are stacked",
label = 'es-catch')
```

\clearpage

## Data and assessment{-}

The last stock assessments for black rockfish were conducted in 2007 for areas north and south of Cape Falcon (45\textdegree46' North latitude). The current assessments assume three areas instead of two, delineated by the state lines as was agreed upon at a pre-assessment and data workshop in March 2015. The prior assessments used Stock Synthesis 2, while the current assessments use Stock Synthesis 
```{r es-modelversion, echo=FALSE} 
if (exists("model")) {
  strsplit(model$SS_version,";")[[1]][1]
}
```
. The Washington base assessment includes a dockside and tag-based CPUE series, but does not include the abundance estimate time series from that same tagging study which was included in the last assessment due to too many violations in the assumptions of abundance estimation. The same two commercial and single recreational fleets are used as in the last assessment for Washington. The Oregon assessment has three commercial fleets and two recreational fleets, while using five surveys and an additional research study for biological compositions. California also has three commercial fleets and 1 recreational fleet with three surveys of abundance, all based on recreational fisheries. All area models include age data as conditional age at lengths. Length compositions are also included in all models.

The stock assessment for `r spp` off Oregon was developed using the length- and age-structured model Stock Synthesis (version 3.30.16).  No previous stock assessment for `r spp` off Oregon has been conducted. Model structure included two fleets (commercial and recreational) and one fishery-based index of abundance. Life history parameters were sex-specific (i.e., a two-sex model) with natural mortality and growth parameters estimated, along with recruitment.  The model covers the years `r startyr` to `r endyr`, with a 12 year forecast beginning in `r endyr + 1`.

This assessment integrates data and information from multiple sources into one modeling framework. Specifically, the assessment uses landings data, length and conditional age-at-length composition data (using ageing error matrices to incorporate ageing imprecision) for each fishery, and one index of abundance based on the recreational fishery; fixed parameterizations of weight-at-length, maturity-at-length, and fecundity-at-length, the Beverton-Holt stock-recruitment steepness value and recruitment variability. Estimated values include initial population scale ($lnR_0$), natural mortality and growth for each sex, asymptotic selectivity and recruitment deviations. The base model was tuned to account for the weighting of the length and age data and index variances (which was estimated), as well as the specification of recruitment variance and recruitment bias adjustments.  Derived quantities include the time series of spawning biomass, age and size structure, and current and projected future stock status. 

Within model uncertainty is explicitly included in this assessment by parameter estimation uncertainty, while among model uncertainty is explored through sensitivity analyses addressing alternative input assumptions such as data treatment and weighting, and model specification sensitivity to the treatment of life history parameters, selectivity, and recruitment. A reference model was selected that best fit the observed data while concomitantly balancing the desire to capture the central tendency across those sources of uncertainty, ensure model realism and tractability, and promote robustness to potential model misspecification.

## Stock biomass and dynamics{-}

Spawning stock outputs are all at or above limit reference points (Table ES-2. Only California shows declines significantly below this reference point at any point in the time series. California and Washington stocks show a declining population through most of the 20th Century, with stronger declines in the 1980s, and recoveries beginning in the mid-1990s. Oregon stocks follow this pattern, but with a decline in the most recent period. California (33%) is below the target biomass reference point with an increasing biomass trend (Figures ES-4 and ES-5).  The Oregon stock dropped after the quick ramp up of catches in the late 1970s and continued a steady decline until around year 2000, settling in at a stock status around 60% of initial conditions. The Washington stock, currently 43%, dropped below the target biomass by in the early 1980s, then risen above since the late 1990s and has fluctuated above that point through 2014 (Figures ES-8 and ES-9).

Spawning output (in millions of eggs; meggs) instead of spawning biomass is used to report the mature population scale because fecundity is nonlinearly related to body female weight. The estimated spawning output at the beginning of 2021 was `r ssb_2021` meggs (~95 percent asymptotic intervals: `r ssb_2021_low` to `r ssb_2021_hi` meggs, Table \ref{tab:ssbES} and Figure \ref{fig:es-ssb}), which when compared to unfished spawning output (`r ssb0`) meggs gives a relative stock status level of `r round(100*depl_2021,0)` percent (~95 percent asymptotic intervals: `r round(100*depl_2021_low,0)` to `r round(100*depl_2021_hi,0)` percent, Figure \ref{fig:es-depl}).  Overall, spawning output declined with the onset of increasing commercial removals in the 1960s and continued to decline with the increase in recreational catches through the 1990s, even dropping below the target relative stock size. The largest of the estimated recruitment pulses since the mid 1990s (that are supported by each of the data sets) caused a sharp increase in spawning output through the mid 2010s, followed by another decline. The minimum relative stock size of `r round(100*as.numeric(depl_min),0)` percent of unfished levels is estimated to have occurred in `r depl_min_yr`. Currently the stock is estimated well above the management target of $SB_{40\%}$ in 2021 and is estimated to have remained above the target since 2000 (Table \ref{tab:ssbES} and Figure \ref{fig:es-depl}).

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots",
"ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
label = 'es-ssb')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts9_Relative_spawning_output_intervals.png"), 
caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
label = 'es-depl')
```

\clearpage

## Recruitment{-}

The California model shows a few extraordinarily high recruitment events that are supported by the length composition data, index data and on-the-water reports (Table ES-3; Figure ES-10).  Oregon recruitment is highly uncertain (Table ES-3; Figure ES11).  Washington recruitment is dynamic, but also shows the most informed recruitment time series, which is consistent with the extent of length and age compositions available to that assessment (Table ES-3; Figure ES12). Both California and Washington support elevated recruitment in the late 2000s.

Informative recruitment begins in the 1960s and peaks in the 1990s (Table \ref{tab:recrES} and Figure \ref{fig:es-recruits}). Data were most informative from the the 1990s to the mid-2010s. Peaks years of recruitments are found in years 1993, 1994, 1998, 2005 and 2015 (Figure \ref{fig:es-rec-devs}). Overall, the `r spp` stock has not been reduced to levels that would provide considerable information on how recruitment compensation changes across spawning biomass levels (i.e., inform the steepness parameter).  Thus, all recruitment is based on a fixed assumption about steepness ($h$ = 0.72) and recruitment variability ($\sigma_R$ = 0.6).

\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
caption = "Estimated time series of age-0 recruits (1000s) for the base model with 95 percent intervals",
label = 'es-recruits')
```

#```{r, results = 'asis'}
#add_figure(
#filein = file.path(mod_loc, "plots", "recdevs2_withbars.png"), 
#caption = "Estimated time series of recruitment deviations",
#label = 'es-rec-devs')
#```

## Exploitation status{-}

California and Washington models indicate that current fishing practices are near or above the SPR rate fishing intensity target, while the Oregon model is quite a bit above the target (table ES-4, compare to SPR=0.5; Figure ES-13 to Figure ES-18), though the steepness value (0.773) indicates a much lower value of SPR for sustainable removals. Fishing rates have been above the target in California in nearly all years since the 1980s, but have dropped considerably in recent years. Oregon fishing rates have been consistently high in recent years.  Washington shows a dramatic decline in fishing intensity since the late 1990s and has fluctuated mostly below the target since. 

Trends in fishing intensity (1 - SPR) largely mirrored that of landings until the 1990s when recruitment pulses overcame the catches to lower overall fishing intensity (Figure \ref{fig:es-1-spr}). The maximum fishing intensity was `r spr_max` in `r spr_max_yr`, above the target SPR-based harvest rate of 0.50 (1 - $\text{SPR}_{50\%}$). Current levels of `r spr_recent_10[10]` for 2020 are near the fishing limit. Fishing intensity over the past decade has ranged between `r min(spr_recent_10)` and `r max(spr_recent_10)` and the exploitation rate has been high (`r min(exploit_recent_10)` - `r max(exploit_recent_10)`, Table \ref{tab:exploitES}). Current estimates indicate that `r spp` spawning output is much greater than than the target biomass level ($\text{SB}_{40\%}$), though fishing intensity remains near target $F_{MSY}$ proxy harvest rate.

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year for the base model. The management target is plotted as a red horizontal line and values above this reflect harvest in excess of the proxy harvest rate",
label = 'es-1-spr')
```

\clearpage


## Ecosystem considerations{-}

This stock assessment does not explicitly incorporate trophic interactions, habitat factors or environmental factors into the assessment model. More predation, diet and habitat work, and mechanistic linkages to environmental conditions would be needed to incorporate these elements into the stock assessment. While no mechanisms have been put forth for these time-varying changes in growth, an environmental component is possible. Limited data in Oregon and California also suggest the possibility that growth has changed over time.

## Reference points{-}

Reference points were based on the rockfish FMSY proxy (SPR50%), target relative biomass (40%) and model-estimated selectivity for each fleet. California is below the target biomass reference point, but above the limit reference biomass (25%). Oregon is well above the target biomass.  Washington relative biomass is above the target biomass. California and Washington yield values are lower than the previous assessment for similar reference points due to lower overall natural mortality values (Table ES-5). The proxy MSY values of management quantities are the most conservative compared to the estimated MSY and MSY relative to 40% biomass for both California and Washington (Table ES-5). The equilibrium estimates of yield relative to biomass are provided in Figure ES-19 to Figure ES-21.

The 2021 spawning biomass relative to unfished equilibrium spawning biomass is well above the management target of 40 percent of unfished spawning biomass. The relative biomass and the ratio of the estimated SPR to the management target ($\text{SPR}_{50\%}$) across all model years are shown in Figure \ref{fig:es-phase} where warmer colors (red) represent early years and colder colors (blue) represent recent years.  There have been periods where the stock status has decreased below the target and fishing intensity has been higher than the target fishing intensity based on $\text{SPR}_{50\%}$. Figure \ref{fig:es-yield} shows the equilibrium curve based on a steepness value fixed at `r round(model$parameters[model$parameters$Label == "SR_BH_steep","Value"], 2)` with vertical dashed lines to indicate the estimate of fraction unfished at the start of `r endyr +1` (current) and the estimated management targets calculated based on the relative target biomass (B target), the SPR target, and the maximum sustainable yield (MSY).

Reference points were calculated using the estimated selectivities and catch distributions among fleets in the most recent year of the model, `r endyr` (Table \ref{tab:e_ReferencePoints_ES}). Sustainable total yield, removals, using an $\text{SPR}_{50\%}$ is `r format(model$derived_quants[model$derived_quants$Label == 'Dead_Catch_SPR', 'Value'], digits = 3, scientific = FALSE, big.mark = ',')` mt. The spawning output equivalent to 40 percent of the unfished spawning output ($\text{SO}_{40\%}$) calculated using the SPR target ($\text{SPR}_{50\%}$) was `r format(model$derived_quants[model$derived_quants$Label == 'SSB_SPR', 'Value'], digits = 4, scientific = FALSE, big.mark = ',')` meggs. Recent removals have been close to the point estimate of potential long-term yields calculated using an $\text{SPR}_{50\%}$ reference point and the population size has been relatively decreasing toward the target over the past few years.

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the base model",
label = 'es-phase')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the base case model. Values are based on the 2020
fishery selectivities and with steepness fixed at 0.80",
label = 'es-yield')
```

\input{tex_tables/e_ReferencePoints_ES.tex}

\clearpage

## Management performance{-}

Removals have been below the equivalent ABC-ACL since the prior assessment (Table ES-6), but those specified ABCs from the 2007 assessments are higher than those coming from the current assessment models. Removals over the last few years have or may have exceeded the newly estimated ABC-ACL values in some years. The differences in the treatment of natural mortality between the previous and current assessments are the biggest reason for this discrepancy.

Exploitation on `r spp` increased starting around 1960 and reached a high in the early 1990s. Since that time, catch has mostly fluctuated between 5 and 10 mt per year, with some years exceeding 10 mt, particularly in the last 4 years. The last ten years of the vermilion rockfish component acceptable biological catch (ABC) and annual catch limit (ACL) (which are equivalent) of the Minor Shelf Rockfish North Complex has been set, by definition, below the overfishing limit (OFL) (Table \ref{tab:ofl-es}). The `r spp` component OFL for this Complex has been exceeded by the Oregon removals in the most recent 4 years.

```{r, results = 'asis'}
yrs = hist = 2011:endyr
nfleets = 2
catch = dead = est.dead = input.catch =  0
for (i in 1:nfleets){
  name = paste0("retain(B):_",i)
  input.catch = model$timeseries[model$timeseries$Yr %in% yrs, name]
  catch = cbind(catch, input.catch)

  name = paste0("dead(B):_",i)
  est.dead = model$timeseries[model$timeseries$Yr %in% yrs, name]
  dead = cbind(dead, est.dead)
}
total.catch <- round(apply(catch, 1, sum),1)
total.dead  <- round(apply(dead, 1, sum), 1)

man = read.csv("C:/Users/Jason.Cope/Documents/Github/Sebastes_melanops_WA/Document/write_up/tables/vermilion_man_for_doc.csv")
man = man[man$Year %in% yrs, ]
out = cbind(man,  total.catch, total.dead)

col_names = c("Year", "OFL", "ABC", "ACL", "Landings", "Est. Total Mortality")

table_format(x = out,
             caption = "The OFL, ABC, ACL, landings, and the estimated total mortality in metric tons.",
             label = "ofl-es",
             align = 'l',
             col_names = col_names)
```

## Unresolved problems and major uncertainties{-}

The most significant uncertainty for all models is the treatment and value of natural mortality and the form of fleet selectivity (e.g., length-based asymptotic vs. age-based dome-shaped selectivity). Data-driven selection between the extreme “kill” (using a ramping of M) or “hide” hypotheses are not currently resolvable. The current California and Washington base models instead use a form of the “kill” hypothesis by not implementing the age-based selectivity (“hide” hypothesis) and estimating female and male natural mortality, thus avoiding a fixing natural mortality as was necessary in the Oregon model. The Oregon model also contained a step in female natural mortality, a specification not used in the California or Washington models. Another important issue is the highly uncertain historical time-series of removals in all states, which needs further consideration. The development of fishery-dependent indices of abundance still requires further attention. Steepness, while fixed, is still highly uncertain for rockfishes and currently is mismatched to the MSY proxy. And while the steepness profile shows low sensitivity in several derived quantities, steepness strongly defines the yield capacity of stocks, and therefore could cause major uncertainty in the recommended management quantities. Stock structure and its relationship to the current political/management boundaries are also not fully understood, both within U.S. jurisdiction and between the U.S. and Canada. While this is a common challenge faced in most west coast stock assessments, further improvement on this topic will likely rely on black rockfish-specific data.

## Scientific uncertainty{-}

The model-estimated uncertainty around the `r endyr+1` spawning biomass was $\sigma$ = `r round(model$Pstar_sigma,2)` and the uncertainty around the OFL was $\sigma$ = `r round(model$OFL_sigm, 2)`. This is likely an underestimate of overall uncertainty because of the necessity to fix some parameters such as steepness, as well as a lack of explicit incorporation of model structural uncertainty.

## Harvest Projections and Decision Table{-}

Black rockfish assessments for California and Washington have a preliminary distinction as category 1 stock assessments, thus harvest projections and decision tables are based on using P*=0.45 and sigma = 0.36, resulting in a multiplier on the OFL of 0.956. The Oregon black rockfish assessment is a category 2 assessment, with a P*=0.45 and sigma = 0.72 with a multiplier of 0.913 applied to the OFL.  These multipliers are also combined with the rockfish MSY proxy of FSPR=50% MSY and the 40-10 harvest control rule to calculate OFLs, ABCs and ACLs. Projections for each state are provided in Table ES-7 to Table ES-9. 

Uncertainty in management quantities for the base model of each state was characterized by exploring various model specifications in a decision table. Initial exploration included natural mortality and steepness values, and uncertainty in historical trawl catches for the WA and CA models. OR explored the scale factor coming from the value of the tagging catchability (Q) parameter, as well as M values. For the CA and WA models, there was very little sensitivity to steepness and trawl catches, but natural mortality produced sensitive results of predicted population scale and status. Discussion with the STAR panel resulted in high and low states of nature +/- 0.03 from the base case natural mortality values for females and males. High and low catch streams (rows) were determined by the forecasts, as described above, for each state of nature. Thus the low catch stream is based on the forecast from the low state of nature. The OR model demonstrated little sensitivity to M, but high sensitivity to the tagging survey Q. High and low states of nature, respectively, were based on a fixed tag of Q = 0.125 and Q estimated by the model. Resultant decision tables are provided in Table ES-10 to Table ES-12.

A ten year (2023-2032) projection of the reference model with removals in 2021 and 2022 provided by the Groundfish Management Team for each fleet under the category 1 (sigma=0.5) time-varying buffer using $P^*$ = 0.45 and 40-10 ABC control rule is provided in Table \ref{tab:project_ES}.

 \input{tex_tables/project_ES.tex}

The decision table (Table \ref{tab:es-dec-tab}) was constructed using female and male natural mortality to define the low and high states of nature. The multi-parameter likelihood profile was used to find the low (Female M = 0.07092; Male M= 0.06525) and high (Female M = 0.08527; Male M = 0.07845) female and male natural mortality values that produce -log likeliehood values  +0.66 units from the reference -log likelihood value. These correspond to the 12.5% and 87.5% quantiles (standard quantiles used in west coast decision tables). The catch rows in the table were based on three proposed catch streams:
  1. P* = 0.45, sigma = 0.5
  2. P* = 0.40, sigma = 0.5
  3. An equilibrium catch based on the $F_{MSY}$ proxy using SPR = 0.5.  

Across all states of natures and catch streams, vermilion rockfish relative stock size never falls below the target relative stock size of 40%. Both P* approaches lower the stock status from the high relative stock size values, while the $F_{MSY}$ proxy does not. The mismatch in the corresponding steepness value ($h=0.6$) that matches MSY at SPR = 0.5 with the steepness value in the stock assessment ($h=0.72$) that correpsonds to an MSY SPR of 0.35 explains why this constant catch will maintain the stock at very high relative stock status levels. 

\clearpage

 \input{tex_tables/decision_table_es.tex}

\clearpage


## Research and data needs{-}

Recommended avenues for research to help improve future black rockfish stock assessments:

1. Further investigation into the movement and behavior of older (> age 10) females to reconcile their absence in fisheries data. If the females are currently inaccessible to fishing gear, can we find where they are?
2. Appropriate natural mortality values for females and males. This will help resolve the extent to which dome-shaped age-based selectivity may be occurring for each.
3. All states need improved historical catch reconstructions. The trawl fishery catches in particular require particular attention. Given the huge historical removals of that fleet in each state, the assessment is very sensitive to the assumed functional form of selectivity. A synoptic catch reconstruction is recommended, where states work together to resolve cross-state catch issues as well as standardize the approach to catch recommendations.
4. Identifying stanzas or periods of uncertainty in the historical catch series will aid in the exploration of catch uncertainty in future assessment sensitivity runs. 
5. The ODFW tagging study off Newport should be continued and expanded to other areas.  To provide better prior information on the spatial distribution of the black rockfish stock, further work should be conducted to map the extent of black rockfish habitat and the densities of black rockfish residing there.
6. An independent nearshore survey should be supported in all states to avoid the reliance on fishery-based CPUE indices.
7. Stock structure for black rockfish is a complicated topic that needs further analysis. How this is determined (e.g., exploitation history, genetics, life history variability, biogeography, etc.) and what this means for management units needs to be further refined. This is a general issue for all nearshore stocks that likely have significant and small scale stock structure among and within states, but limited data collections to support small-scale management.


<!--chapter:end:01executive.Rmd-->

\vspace{500cm}

<!--# Disclaimer{-}

_**These materials do not constitute a formal publication and are for information only. They are in a pre-review, pre-decisional state and should not be formally cited or reproduced. They are to be considered provisional and do not represent any determination or policy of NOAA or the Department of Commerce.**_-->

\pagebreak
\setlength{\parskip}{5mm plus1mm minus1mm}
\pagenumbering{arabic}
\setcounter{page}{1}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}

<!--chapter:end:10a.Rmd-->

# Introduction
## Basic Information
`r Spp` (*Sebastes melanops*) are an important component of the recreational fisheries in the nearshore waters off central and northern California, Oregon, and Washington, as well as the non-trawl commercial fisheries in California and Oregon. They range as far north as Amchitka and Kodiak islands in Alaska and are considered uncommon south of central California [@love_rockfishes_2002].  

A first assessment of `r spp` off considered the population off Oregon and California [@ralston_status_2003] and reviewed the evidence supporting genetic stock structure for `r spp` and other rockfish off the U.S. West Coast and concluded that the Oregon and California populations of `r spp` are probably not genetically heterogeneous.  That assessment treated the `r spp` off California and Oregon as a unit stock.  Previous assessments of `r spp` off Washington [@wallace_status_1999; @wallace_status_2007] describe a study of coastal `r spp` genetic structure using 10 sampled sites collected from northern California to southern British Columbia t 1995-97.  Results of that study support the notion of separate genetic stocks north and south of Cape Falcon.  However, a later study [@baker_genetic_1999] of `r spp` collected from eight sites along the northern Oregon coast concluded that `r spp` from north and south of Cape Falcon were genetically very similar.

Although a stock boundary line at the Columbia River seems reasonable for `r spp`, both because it is a state fishery management boundary and because the Columbia River plume is likely to be a natural barrier to the north-south exchange of `r spp` adults and larvae, the 2007 assessment of `r spp` off Oregon and California [@sampson_status_2007] differed slightly from Ralston and Dick [-@ralston_status_2003] in placing the northern boundary at Cape Falcon rather than at the Columbia River.  The boundary was changed to avoid overlap with the separate northern assessment [@wallace_status_2007] and to simplify the process of assembling historical commercial landings data, which are largely available in terms of Pacific Marine Fisheries Commission (PMFC) statistical areas.  The northern boundary of PMFC Area 2C is at Cape Falcon (Figure 1).  Given the spatial resolution of the historical commercial fishery data, it is very problematic to estimate the catch of `r spp` taken north of Cape Falcon but south of the Columbia River.

During a preliminary workshop in April 2015 [@pfmc_safe_2015], to discuss approaches for assessing `r spp`, China rockfish (*S. nebulosus*), and kelp greenling (*Hexagrammos decagrammus*), it was agreed that the assessments for these nearshore species should at a minimum be spatially stratified with boundaries at the CA/OR border (42\textdegree00' N latitude) and the OR/WA border (46\textdegree16' N latitude).  Such a spatial stratification would be consistent with two ideas: (a) these nearshore species do not exhibit much adult movement and (b) exploitation and management histories have varied significantly among the three states.  Together these features would likely create appreciable state-to-state differences in age composition for each of the three species.  Stock assessment teams were advised that they could use geographic strata that were finer than the state level if there were data to support such an approach (Figure 1).

At the same nearshore stock assessment workshop, it was agreed that recreational catch histories for the stocks of `r spp` should be assembled on the basis of port of landing rather than location of fish capture, even though fishing vessels landing their catches into a port in one state might have captured fish in waters off a neighboring state.

Accounting for location of capture is very problematic for recreationally caught fish and for commercial catches taken with non-trawl types of gear (e.g., hook-and-line), for which there are no or very limited logbooks that report fishing location.  For these regional assessments the commercially caught `r spp` were apportioned to assessment region based on the port of landing, with the exception of trawl caught fish landed into Astoria, OR.  Most of these fish were assumed to have been caught off Washington and most of the trawl landings into Astoria were therefore included with the catch history for the Washington assessment region.  Details are provided below in Section 2.1.1.1 The PacFIN Era (1981 to 2014).

## Life History
Adults tend to occur in schools over rocky structure at depths less than 40 fathoms, and sometimes feed actively on or near the surface.  They feed on a wide variety of prey including zooplankton, krill, mysids, sand lance, and juvenile rockfish, and are subject to predation by lingcod and marine mammals [@love_rockfishes_2002]. 

Although tagging studies have documented some individuals moving long distances (several hundreds of miles), the vast majority of recaptured individuals were found close to the areas of initial capture and tagging [@culver_1987;@ayres_tag_1988;@wallace_tag_2010;@starr_green_2007].  Results from a 2004-05 study off Newport, OR of 42 `r spp` implanted with acoustic tags indicated that all but seven fish remained within range of a 3 x 5 km array of acoustic receivers during one full year of monitoring and had relatively small home ranges that did not vary seasonally [@parker_etal_2007].  Green and Starr [-@green_starr_2007] report similar findings from a study in Carmel Bay, CA of 23 acoustically tagged `r spp`. The extensive Washington state tagging study also supported low movements for most individuals, with some exceptional movements recorded [@wallace_tag_2010].

Like all members of the genus Sebastes, `r spp` have internal fertilization and bear live young approximately two months after insemination. `r spp` are quite fecund, with a six-year-old female annually producing about 300,000 embryos and a 16-year-old producing about 950,000 embryos [@bobko_berkeley_2004].  Recent studies have demonstrated that the relative number and quality of larvae increase with age in female `r spp` [@berkeley_etal_2004; @hixon_etal_1987]. Parturition of larvae occurs during winter [@wyllie_echeverria_1987] and larvae and small juveniles are pelagic for several months to a year [@boehlert_yoklavich_1983].  Settlement occurs in estuaries, tide-pools, and in the nearshore at depths less than 20 m [@stein_hassler_1989].  

`r spp` begin recruiting to nearshore fisheries at 3-4 years of age, corresponding to a fork length of about 25-30 cm, and 50% of females attain maturity at about 6-8 years, corresponding to a fork length of about 38-42 cm.  Adult female `r spp` grow 3-5 cm larger than males, with a few females attaining fork lengths greater than 55 cm.

## Ecosystem Considerations
No formal ecosystem considerations have been made given the lack of data for such an undertaking. Differences in growth though time have been considered in the model specification in the Washington model. Though the mechanism is not specified, this could certainly be due to process error driven by environmental conditions.

## Historical and Current Fishery Information
`r spp` are harvested by a wide variety of fishing methods including trawling, trolling, and hook-and-line fishing with jigs and long-lines.  Although `r spp` have never been a dominant component of any commercial fisheries, they are important as incidental catch in the troll fishery for salmon and the troll and jig fisheries for groundfish.  With the decline of salmon fishing opportunities in the late 1970s and early 1980s `r spp` became a vital target of marine recreational fisheries in Oregon and Washington, especially during periods of restricted or slack fishing for salmon, halibut, and tuna.  

`r spp` are also an important component of the recreational fisheries in northern California but are of less significance south of Cape Mendocino due to their reduced prevalence compared to other species. Since 1990 annual recreational harvests of `r spp` have averaged 229.6 tons off California, 304.4 tons off Oregon, and 272.5 tons off Washington. Commercial annual harvests by non-trawl gear types during the same period averaged 44.6 tons in California, 62.0 tons in Oregon, and 14.7 tons in Washington. Harvests by trawl on average during this period have been less than 19.3 tons annually for all three states combined.

Removal histories have been a significant axis of uncertainty in the past assessments of `r spp`.  Because of concerns about the effects of initial equilibrium assumptions on the level of depletion estimated in the preliminary base model, the 2003 Stock Assessment Review (STAR) panel worked with the Stock Assessment Team (STAT) to develop a catch history that avoided the need to assume historical catch and equilibrium conditions in the first year of the assessment.  The assumed catch reconstruction began in 1946, ramping up from zero in 1945 and all prior years.  In hindsight, this may not have been a good assumption, as indicated by the following text from [@cleaver_1951] that describes catches of rockfish from 1941 to 1949 in Oregon.

"The rockfish are caught by otter trawl and long-line gear.  The principal species caught by the otter trawl are the `r spp` (*Sebastodes melanops*); green or yellowtail rockfish (*S. flavidus*); red or orange rockfish (*S. pinniger*); and rosefish (*S. alutus*). The landings of rockfish (all species) rose rapidly during the war from 1,301,400 pounds in 1941 to a peak of over 17,000,000 in 1945. Subsequently the landings fell rapidly because of decreased demand and leveled off at about 4,000,000 per year in 1949.

Cleaver [-@cleaver_1951] also states, in an introductory section on Bottom Fisheries, that the "otter trawl fishery accounts for at least 95 percent by weight of the bottom fish landings."

That `r spp` is one of only four species that Cleaver [-@cleaver_1951] identifies as composing the large landings of rockfish in Oregon (most of which was actually taken off of Washington waters) during WWII suggests that `r spp` were not a trivial fraction of the large catches taken during the 1940s.  One might also suppose that the otter trawl fishery took a large portion of the landings of `r spp`.  Cleaver's statements are certainly at odds with the catch reconstruction developed in the previous assessments.

It seems that `r spp` were also landed in appreciable quantities in California during the 1940s.  `r spp` was identified by scientific name as one of the "half-dozen of the larger and more abundant species [that] make up over half of the annual California commercial poundage landed" (Anon. 1949).

A major task for the 2007 assessments of `r spp` in was developing a plausible reconstruction of historical landings of `r spp` and exploring the consequences of those landings.  For the current set of assessments catch histories from the past assessments have been reconsidered. Formal catch reconstructions have been conducted in California [@ralston_documentation_2010] and Oregon [@karnowski_historical_2014], but even those relatively newer attempts were reconsidered in light of contributions from state agencies. For this assessment, Washington provided a first step in an approach to provide a reconstructed historical catch time series for a stock, something needed for all species in the state's waters.

## Summary of Management History and Performance
Prior to 2000 the Pacific Fishery Management Council (PFMC or Council) managed the fishery for `r spp` as part of the Sebastes complex, with no separate Acceptable Biological Catch (ABC) or Optimum Yield (OY) for `r spp`.  In 2000 the Council established an ABC of 1,200 mt for `r spp` caught north of Cape Mendocino (in the Eureka, Columbia, and Vancouver INPFC statistical areas), but left `r spp` south of Cape Mendocino as part of the "other rockfish" category.  For 2001 through 2003 the ABC for `r spp` caught north of Cape Mendocino was 1,115 mt annually, and `r spp` south of Cape Mendocino remained part of the "other rockfish" category and without a separate ABC or OY.

Regulation of the `r spp` fisheries by the PFMC prior to 2004 was accomplished primarily by trip limits for commercial fisheries and bag limit restrictions for recreational fisheries, with different limits applying in different geographic regions (see Table 1 in @ralston_status_2003).  Some other important regulations include the following.

* 1953: California prohibited trawling within three miles of shore.
* 1995: The commercial hook-and-line fishing in Washington state waters (0-3 miles) was closed to preserve recreational fishing opportunities and avoid localized depletion; the closure was extended to trawlers in 1999.  Oregon established `r spp` management areas with reduced daily commercial fishery trip limits in area near ports with large recreational fisheries.
* 2000: `r spp` began to be managed by the Council as a minor nearshore species.  Commercial trip limits were significantly reduced, with specific restrictions applying to `r spp`.  California instituted seasonal closures for commercial and recreational fisheries inside 20 fathoms, reduced the bag limit for rockfish from 15 to 10 fish, and limited recreational gear to one line with three hooks.
* 2002: California adopted a Nearshore Fishery Management Plan and began more active management of nearshore fisheries including the use of seasonal, regional, and depth-specific closures.  Oregon adopted an Interim Nearshore Fishery Management Plan in anticipation of increased pressure on nearshore stocks due to reduced fishing opportunities for groundfish in federal waters.  Regulations included fishing-sector specific caps on retained harvests, set approximately at the levels attained in 2000.
* 2003: The Council established Rockfish Conservation Areas (RCAs) to control catches of overfished rockfish species, and large portions of the shelf were closed to fishing.  Differential trip limits were applied north and south of a management boundary at 40\textdegree10' N. latitude for nearshore Sebastes species. Nearshore permittees in California became subject to depth restrictions consistent with the shoreward non-trawl RCA boundary.  In California the commercial and recreational fisheries for rockfish were closed early.
* In 2004 and 2005: the sport fishery in Oregon closed in September 2004 due to early attainment of the state's limit for sport-caught `r spp`.  This was the first time that the sport rockfish fishery in Oregon had not been open all year.  In 2005 it closed early again.
* In 2008 the groundfish trawl fishery was closed in Washington from the seaward RCA boundary to the shore north of 48\textdegree10' N. latitude to address increased encounters with yelloweye rockfish and canary rockfish.

In recent years California, Oregon and Washington regulations for the marine sport fisheries, which has been the major source of mortality on `r spp`, have become quite complicated and variable through time.  Tools for regulating the sport fishery include closed areas, depth restrictions, seasonal closures, and bag limits.

California had no recreational bag limit for rockfish until 1990 when a 15 fish per day per angler limit was implemented.  In 2000 the limit was reduced to 10 fish per day for each angler's combined bag of rockfish, cabezon and greenling.  The fishing season was year-round prior to 2000 and since then has been variable by state management area.  There were no gear restrictions prior to 2000.  In 2000 anglers were limited to fishing one line with three hooks.  Since 2001 they have been restricted to one line with two hooks.  There is no minimum size limit for `r spp`.

Oregon had no recreational bag limits for marine fishes until 1976 when the state established a 25-fish limit.  In 1978 the state established a daily limit of 15 fish for each angler's combined bag of rockfish, cabezon and greenling, which stayed in effect until 1994 when the state established a 10-fish-per-angler daily bag limit specifically for `r spp`.  Following the early closure of the fishing season for `r spp` in 2004, the daily bag limit for `r spp` was dropped to 5 fish at the start of 2005 but was increased in-season to 6 fish.  The per-angler daily bag limit was 6 fish during 2006 and 2007, 5 fish at the start of 2008 and increased in-season to 6 fish, 6 fish at the start of 2009 and increased in-season to 7 fish where it has remained since.  

The goal of Oregon's sport fishery management is to maintain year-round fishing opportunities.  In-season adjustments to regulations can be made more restrictive or less restrictive, depending on circumstances and the prospects for early attainment of harvest caps.  Seasonal depth restrictions (e.g., inside 30 fathoms April 1 to September 30) are one tool used regularly in recent years to control the fishery, driven largely by the need to avoid bycatch of the primary rebuilding species, canary rockfish and yelloweye rockfish.

Washington had a recreational daily bag limit for rockfish (all species) of 15 fish per day from 1961 to 1991, 12 fish per day from 1992 to 1994, and 10 fish per day from 1995 to 2015.  The bag limit for blue rockfish plus `r spp` in Marine Area 4B (Neah Bay) has been 6 fish per day since 2010.  Fishing seasons for groundfish species are structured to provide year-round fishing opportunities, if possible.  Depth restrictions vary by state management area, being more restrictive in the north compared to the south due to higher encounter rates with overfished yelloweye rockfish and canary rockfish.  There is no minimum size limit for `r spp`.

In 2004 the coastwide ABC established for `r spp` was based on the projected yields derived from separate northern [wallace_status_1999] and southern [ralston_status_2003] stock assessments (Table 1). The northern assessment covered the Washington coast and the northernmost portion of Oregon, from Cape Falcon to the WA/OR border at the Columbia River.  The southern assessment covered the entire Oregon coast and the California coastline north of Point Arena.  

To account for the spatial overlap of the two assessment areas, 12% of the projected yield from the northern assessment was transferred to the southern region when deriving the coastwide ABC and OY values of 1,315 mt for 2004.  State-by-state harvest guidelines were established: 326 mt for California, 450 mt for Oregon, and 540 mt for Washington.  A similar approach was taken in 2005 and 2006 and the OY for the area south of the Columbia River was apportioned to harvest guidelines for California and Oregon based on a 42:58 split.  The basis for this apportionment is unclear was to support separate harvest guidelines for each state. The catches were apportioned by the average catch share by state in the 1985-2002 period [@pfmc_abc_2004]. 

In all years when there has been an OY specified for `r spp` the estimated catch has been less than the OY, except for 2003 when the estimated coastwide catch exceeded the ABC for north of Cape Mendocino.  In 2003 the estimated coastwide catch exceeded the OY by 183 mt for the region north of Cape Mendocino, but 290 mt of this coastwide catch was recreational harvest taken south of Cape Mendocino.

## Canadian and Alaska fisheries
`r spp` is a "Non-Quota" species in the Department of Fisheries and Oceans Management Plan, and is not formally assessed in nearshore Canada waters [@dfo_fmp_2014].

Add Alaska assessment text here.

<!--chapter:end:11introduction.Rmd-->

# Data and Model Inputs
A description of each data source is provided below (Figure \ref{fig:data-plot}).

<!--chapter:end:20data.Rmd-->

## Fishery-Dependent Data

### Commercial

#### Landings

The systems along the U.S. West Coast for monitoring commercial fishery landings in the past did not keep track of the landings of individual rockfish species, largely because many rockfish species have similar market characteristics and therefore were landed as an unsorted mix of species.  Black rockfish in particular, which are a nearshore species and much less abundant than many of the offshore rockfish species, were generally landed in mixed-species categories and not required to sort into its own market category until 2006.  As a consequence, the historical records do not provide a detailed accounting of the landings of black rockfish.  The basic approach taken to develop the landings series in this assessment (as in past assessments) was to apply values for the proportion of black rockfish sampled in mixed-rockfish landings.  Data on the proportions of black rockfish are sparse, with the consequence that the landings reconstructions are highly uncertain.


Commercial fishery landings of black rockfish in Washington are compiled from a variety of sources including PacFIN, agency reports, historical fish ticket information and communication with agency personnel (Table 48).  Since 1935, commercial fishing vessels have been required to submit a fish receiving ticket ("fish ticket") for each landing.  Rockfish landings from domestic fishers are usually reported in mixed-species market categories, and are routinely sampled for species composition by port samplers.  The information required on the ticket and sampling methods have changed through time. Due to these changes, we separated the data into three time periods 1935-1969, 1970-1980, and 1981-present based on the level of detail available in the data for compiling landing history in Washington for this assessment. Comparisons of the commercial catch in each fishery for the current and previous assessments (Figure X).



##### Years 1935 to 1969

Although the original paper fish tickets are no longer available for this period, WDFW recently digitized the daily aggregated data from printed reports, to assist in reconstructing the commercial groundfish fishery landings history for Washington.  These daily aggregated data reports contain summaries of daily catches for port-groups by gear and area fished.  The data are available for 1935 and for 1949-1969, and were used as the basis for the black rockfish catch reconstruction for this time period. 

During this period, mixed-species, nominal market categories were typically used for reporting of the aggregated data.  Market categories such as "red rockfish", "black rockfish" (BLK1), and "unidentified rockfish" (URCK) are typical on fish tickets during this time, lumping all red colored fish and black colored fish into these categories for reporting.  For bottom trawl gears, the BLK1 market category consisted of mostly yellowtail rockfish and silvergrey rockfish (Greg Lippert, pers. comm.).  To split the black rockfish landings out of the BLK1 market category, we assumed 10% of the BLK1 landings were S. melanops (BLCK). We further assumed that no other nominal market categories in the trawl fishery contained S. melanops (see table below). 

For the commercial jig and troll fisheries, rockfish were landed in the unidentified rockfish (URCK) market category.  No species composition samples are available during this time, so we assumed 85% of URCK landings were S. melanops, which matches the species composition data from the 1985-1989 commercial jig fishery.  These estimates were also supported by interviews with port samplers active during portions of this period.  The rockfish caught by troll gears composed of mostly yellowtail and black rockfishes.  Wright [-@wright_1967] reported rockfish species composition of the troll landings by port.  We assumed 80% of URCK caught off central Washington were BLCK and 20% for northern Washington landings.  


#####  Years 1970 to 1980

Original fish ticket data were used for commercial catch estimate during this period.  Fishing areas were better defined and reported during this period; there were no longer interstate areas due to the introduction of current management areas, with a boundary line at the OR/WA state border.  However, issues with URCK market category remained.  Species composition sampling of URCK were conducted for trawl and jig fisheries but not for salmon troll and the "other gears" gear types. 

To estimate the trawl landings of black rockfish in the category URCK, we applied the current WDFW species composition algorithm by gear, port, and quarter.  If no species composition samples were taken during a quarter, we first borrowed annual composition data for the gear/port group.  If those data were not available, a coastwide annual composition for the particular gear type was applied.  There was no borrowing of composition information across gear groups or years.

The commercial jig fleet operates in nearshore waters and black rockfish is the dominant component of the URCK landing for this gear type.  Species composition sampling was not conducted during the 1970-1980 time period.  Based on the samples collected in the 1980s, we assumed that black rockfish made up 80% of the total rockfish landed by the jig fleet.  For the troll fishery, the same proportions as for the pre-1970 time period were applied.

Rockfish (URCK) were also landed in small amounts by other commercial fleets, such as fixed gears and salmon troll.  The fleets in the 70s and 80s predominantly targeted sablefish and halibut in waters too deep for black rockfish.  Port samplers did not recall observing any black rockfish in the fixed gear landings (James Beam and Greg Lippert, pers. comm.).  Therefore, we assumed fixed gear landings were negligible.  For URCK landed by the salmon troll fleet, the majority of troll landings was yellowtail rockfish with smaller numbers of widow, canary, and black rockfish (Wendy Beeghley and James Beam, pers. comm.). We assumed that 10% of the troll rockfish landings were black rockfish for 1970-1980.  


##### Years 1981 to the Present (the PacFIN era)

Commercial landings data for the modern era (1981 to present) are taken from the Pacific Fishery Information system (PacFIN), which is a central repository for U.S. West Coast groundfish landings and auxiliary information collected by the three state fishery agencies and other agencies.  A description of basic state data collection systems and overview of PacFIN is provided in Sampson and Crone (1997).  Updated dockside sampling protocols and data processing procedures for Washington are described in Tsou et al. [-@tsou_etal_2015] and Tsou and Weyland [-@tsou_weyland_2015].  Rockfish landings from this period are pulled from the PacFIN comprehensive catch table.  Landings in this table are the products of nominal landings, as well as area and species compositions.  For the remaining URCK in this table, we applied a coastwide annual composition for each gear, as described above.  After this step, there are still small amounts of URCK for trawl and setline gears.  These landings are not included in this assessment. 

For the jig-gear fishery, dockside sampling was conducted by the WDFW Ocean Sampling Program (OSP) during 1985-1991. These species composition data are not available in PacFIN. For landings during 1985-1991 the URCK species compositions were stratified by year and port. For other years, species compositions were stratified by port only.  For jig-caught fish landed into Seattle annual species compositions from Neah Bay were applied (Table 49) because there was no port sampling in Seattle.
The URCK market category was used until 2000, after which it was replaced by the Slope, Shelf, and Nearshore rockfish market categories.  The commercial nearshore fishery was closed in 1999, so starting in 2000 there are negligible black rockfish landings in any market categories. In 2005 mandatory sorting was established for black rockfish, so all black rockfish landed should be recorded on fish tickets in the BLK1 category.

To assign URCK commercial salmon troll landings, we used the same reasoning assumption as applied to earlier periods (see 1970-1980) to assign 10% of the URCK landed in the salmon troll fishery to black rockfish.  After a complete nearshore closure in 1999, black rockfish landings have been negligible.



It has been and continues to be a common practice for Oregon fleets to fish off Washington coast and land the catches at Oregon ports. Although the separate geographic assessments by state region would ideally have strict geographic separation of landed catch to the location of capture, this is not possible to accomplish because information on the fishing location is generally unavailable.  Until recently logbooks that report area of capture were only available for landings by the groundfish trawl fleet.  Oregon has required a logbook for commercial vessels participating in its nearshore fishery since 2004.  

Staff from the Oregon Department of Fish and Wildlife (ODFW) used species composition samples collected during 1976 to 1993 to conduct an analysis of the spatial distribution of black rockfish landed at Astoria, OR.  Astoria is the northernmost port in Oregon and is located near the mouth of the Columbia River, which forms the boundary between Oregon and Washington.  The portion of aggregated rockfish landed pounds that were taken north of the Columbia River (i.e., from waters off Washington) was 98.6%.  This percentage was applied to all historical trawl landings of rockfish at Oregon's Columbia River District ports prior to 1976. It was assumed 14.1% of the aggregated rockfish being Black in the 2015 assessment (Cope et at. 2016). This percentage is much higher than the values used for Washington catch reconstruction. Therefore, we recalculated the portion by using Washington historical species composition data not available in PacFIN for pre-1981 and data available in PacFIN for 1981 to 1986. The revised proportions of black rockfish in the aggregated rockfish landings are 3% and 4% for pre-1981 and 1981-1986, respectively.  Non-trawl landings into Astoria were assumed to have been caught from Oregon waters, same assumption as used in 2015 assessment.  

Starting in 1994 black rockfish landed into Oregon were legally required to be sorted and sold in a separate black rockfish market category and were also reported as separate retained catches in the mandatory trawl logbooks.  Based on the retained catches reported in the logbooks, the estimated proportion of the trawl-caught black rockfish that were caught from off Washington and landed into Astoria ranged from 65 to 100 percent. These black rockfish are accounted for in the Washington regional assessment (Table X).

The Washington Department of Fish and Wildlife (WDFW) provided commercial fishery landings based on fish ticket records of black rockfish harvested off Oregon by vessels landing at ports in Washington.  Landings were less than 1 mt per year for the period 1971 to 2014; therefore, all landings to Washington ports were assumed to occur in waters off Washington in this assessment.



#### Foreign Fishery Catches of Black Rockfish

Rogers [-@rogers_species_2003] developed catch reconstructions for removals by foreign trawlers operating off the U.S. West Coast during the late 1960s to mid-1970s. Although this study reports that Japanese vessels operating in the Columbia and Eureka statistical areas (Oregon and northern California) caught substantial amounts of black rockfish, with cumulative catches of more than 500 mt over 10 years, it seems very unlikely that foreign vessels could have operated sufficiently close to shore to catch appreciable amounts of black rockfish. This assessment does not include Rogers' estimates of foreign fleet removals of black rockfish. 


### Recreational Removals

Comparisons of the catch in each recreational fishery for the current and previous assessments are in Figure X.



The Washington recreational catch history of black rockfish was reconstructed using several direct and indirect records of black rockfish catch (Table X). All primary sources report catch in numbers of fish. As sources have been modified and re-evaluated, a completely new catch reconstruction for Washington was developed for 2015 assessment. This catch history was updated to include 2015-2022 estimates in this assessment.


### Estimated Discards

In the previous assessment, commercial discards were not accounted for due to the information provided by the West Coast Groundfish Observer Program (WCGOP) at that time, showing about a 1% discard rate in their survey.  We evaluated the WCGOP estimates of black rockfish discards from 2002-2013, which showed a total of 32.2 mt in estimated discards and total landings of 2042.5 mt coastwide, resulting in a rough discard rate estimate of 1.58%.  This amount of discards was included in the CA HKL non-live fishery landings, going back to 1916, and in the OR HKL non-live fishery landings to 1892. Given the minimal amount of discards, no further depth-dependent mortality estimates were evaluated for Oregon and California and this discard rate was added to the total commercial removals.  Parker et al [-@parker_buoyancy_2006] concluded that semi-pelagic, vertically mobile species, such as black rockfish, show less barotrauma; hence these estimates could be slightly overestimated.  

Washington recreational discard estimates were not available until 2002.  Numbers of discarded-by-depth black rockfish were estimated using the same catch expansion algorithm for landed catch.  Surface release mortalities adopted by the Groundfish Management Team (GMT) in their death-by-depth matrix (11%, 20%, 29%, and 63%) were then applied to the number of release black rockfish for each of 4 release depth bins (0-10 fm, 11-20 fm, 21-30 fm, and >30fm), respectively. Total dead released black rockfish were then summed across each depth bin. The average weights of discards were assumed to be the same as the average weights of landed and multiplied by the number of released dead to get total dead in metric tons.   For pre-2002 release, proportions of releases based on a ratio estimator using 2003-2007 data were applied.  For the split between charter and private vessels, the same algorithm used for splitting retained catch was applied. The overall average discard rate in the recreational fisheries was 1.37%. There was no information on Washington commercial discards, so the rate of 1.37% (same as the historical recreational discards) was also applied to the entire commercial time series. This low rate was similar to discard rates estimated in the other states.


## Size and Age Composition Data

Fish length measurements, primarily from the recreational fishery, are one of the major sources of data for this assessment.  



### Length and Age Sample Sizes

The level of commercial fishery sampling for black rockfish has been erratic, with almost no samples taken in Washington until the early 1990s (Figure X).  


#### Multinomial Sample Sizes

Initial input values for the multinomial samples sizes determine the relative weights applied in fitting the annual composition data within the set of observations for each fishery.  The initial input values in this assessment were based on the following equation developed by I. Stewart and S. Miller (NWFSC), and presented at the 2006 Stock Assessment Data and Modeling workshop.
Neff = Ntrips + 0.138 * Nfish if Nfish/Ntrips < 44
Neff = 7.06 * Ntrips if Nfish/Ntrips ? 44

Tuning of the assessment model involved multiplying the input sample sizes for each fishery by an adjustment factor to achieve a better balance between how well the model fit the set of composition data and how well it should have fit the data given the sample sizes underlying the data.


#### Length Compositions

The length data for the assessment model were tabulated into 2-cm length bins ranging from 10 cm to 64 cm, with accumulator bins at each end. 

The length composition data indicate some general differences between the three fishery types, with the trawl fisheries producing the largest fish, the recreational fisheries producing the smallest fish, and the non-trawl fisheries producing fish of intermediate length.  There is little evidence in any of the length composition data of distinct modes or successions of modes from one year to the next that might represent strong year-classes.



##### Washington Length Compositions

###### Commercial

The biological data for the commercial fishery were extracted from PacFIN on April 24th, 2015.  These data are from trawl and non-trawl (hook-and-line) fisheries; there is no live-fish fishery off Washington.  Of 9009 records (each representing a single specimen), 4989 were from the trawl fishery (Table X).

Length composition data are reported either in fork length or total length.  Fork lengths are preferred; where they are missing the total length is used. These data are expanded to reduce the effect of non-uniform sampling effort (Table X). The expansions are by weight, catch/sampled catch; first on a per-trip level, and then on a per-year, per-fishery level.  Expansion factors have a minimum value of 1, and are capped at their 90th percentile value. The final sample size is the product of the two expansion factors, which is then capped at its 90th percentile value.  The final sample sizes for the WA biological data ranged from 1 to 389.7, with a median value of 10.4.  

The data were stratified by gender and fishery.  The final sample sizes were stratified and summed by length bin (10 cm to 64 cm bins, 2 cm in width), and an effective sample size is computed from the number of trips and number of fish each stratum represents, according to the Stewart and Miller method for multinomial fishery data. Unsexed fish were treated as above, but entered as a separate dataset.


###### Recreational

The Washington Department of Fish and Wildlife biological database provided sampled length data from the recreational fishery for sexed and unsexed samples for years 1979-2014. Sexed samples were the largest sample sizes and covered most years (Table 53). Composition data were used as collected (i.e., not expanded). Effective sample sizes were based on unique "sequence" sizes, which is roughly equivalent to a trip. 


##### Tagging data

The Washington Department of Fish and Wildlife provided sampled length data from the tagging survey for sexed and unsexed samples for years 1981-2014 (Table 54). Unsexed and sexed data were generally available in different years. Like the recreational data, composition data were used as collected (i.e., not expanded) and effective sample sizes were based on unique "sequence" sizes, which is roughly equivalent to a trip. 


#### Age Compositions

Commercial age composition data were a subset of the length data, 7984 records in total, and were expanded in the same manner as the lengths (Table 55).  Ages were stratified by fishery and gender, and binned in 1-year bins from 0 to 45, with additional bins 50 and 55.  As for the length compositions, the unsexed fish (29 samples) were treated as a separate dataset.  Samples were also available by sex for several years in the recreational data (Table 56).

Age-at-length compositions were not expanded; the final sample sizes were the sum of all individual age samples per length bin. For all three models, the ages were modeled as conditional age-at-length.


### Abundance Indices

Age and length composition data by themselves do not provide sufficient information to reliably determine trends in stock abundance and biomass.  Most assessments of U.S. West Coast groundfish stocks rely on estimates of stock biomass from research trawl surveys to provide information on biomass trends, but black rockfish are very infrequently caught in any of the bottom trawl surveys, which have a limited coverage of shallow nearshore waters (none of the surveys have ever been conducted in waters shallower than 55 m).  

#### Dockside catch-per-unit-effort for Washington

The WDFW provided recreational dockside fisheries data from 1981 through 2022 for consideration in this assessment. Data were collected at the trip level, with the number of landed fish and the number of anglers on each vessel being recorded. The amount of time fished by each angler was not recorded. This data had also been provided for the 2015 Black Rockfish Assessment where authors used the Stephens-MacCall method [@stephens_multispecies_2004] as an objective approach for identifying trip records of catch/effort data, and several covariates including year, month, boat type, area, daily bag limits and depth restrictions. For the 2015 Assessment data were modeled using a delta-GLM approach, where the catch occurrence (binomial) component was modeled using a logit link function and the positive catch component was modeled using either lognormal or gamma distributions. 

The recreational dockside data was considered for use in this assessment, however, management measures implemented over the past several decades impeded extracting a reliable signal for use as abundance indices. In 2003, management restricted summer fishing depths to shallower than 20-fathoms in WDFW marine areas 3 and 4, and in 2006 modified this depth restriction to 30-fathoms in marine areas 2, 3, and 4. Additionally, daily rockfish limits were 15 fish from 1981-1991, 12 fish from 1992-1994 (except in area 1 where it remained at 15), 10 fish from 1995-2016, and a reduction to 7 fish from 2017-2022. The effects of these management changes to CPUE of black rockfish could not be reconciled with changes in the abundance indices and therefore this dataset was not used in this assessment. 

<!--chapter:end:21f-.Rmd-->

## Fishery-Independent Data
### Abundance Indices
#### Nearshore Survey CPUE for Washington
Fishery-independent data available for this assessment came from two distinct WDFW research projects. The first was the Black Rockfish Tagging Program that was initiated in the early 1980s and provided CPUE of black rockfish captured for tag releases primarily off the central coast of Washington. The second data set was from the standardized Coastwide Rod-and-Reel Survey that began in 2019.  While technically independent from one another, these two fishery-independent data sources had comparable components allowing the data to be considered in two different ways. 

The Washington Black Rockfish Tagging Program was initiated in 1981 with the primary objective of collecting biological information such as growth and movement. The program continued through the 1980’s with modifications to protocols including scope, primary objectives, and tagging methods. Details of this extensive program can be found in Wallace et al. [-@wallace_tag_2010].  Beginning in 1998, the geographic range of sampling was constrained, and effort was primarily focused on rocky habitats during spring months off the central coast of Washington in Marine Catch Area 2 (MCA 2).  Sampling crews consisted of 8-15 anglers using rod-and-reel rigged with one to three single hook jigs per line. During the sampling process, catches of black rockfish per angler minute were recorded, as were covariates month and MCA. Black rockfish were targeted during each trip. In 2010, the WDFW expanded the tagging program to include additional nearshore bottomfish species and increased the geographic disbursement of tags to the entire 180-mile-long Washington coastline. The program retained a primary objective of targeting black rockfish during the spring tagging efforts.

Tag release data collected from the Tagging Program were used for constructing abundance indices in all previous assessments for Black Rockfish off Washington coast. The 2009 stock assessment review panel considered tagging q as one of the major uncertainties due to the spatial coverage of the tag-release sites focusing mainly on central Washington coastal waters.  The 2015 Stock Assessment Review panel recommended future research to include definition and measurement of black rockfish habitat, the development of a coastwide fishery-independent survey for nearshore stocks, and improving CPUE standardization protocols [@brf_STARrev_2015]].

Beginning in 2010, the WDFW started to address STAR panel comments and recommendations. In 2011, geographic coverage of the Tagging Program was expanded by adding more stations to the northern and southern coastal waters, while black rockfish remained the targeted species. In 2014, the WDFW decided to terminate the historical Black Rockfish Tagging Program and started to plan for a survey to include other nearshore groundfish species besides black rockfish, such as China, quillback, copper, lingcod, cabezon, and kelp greenling.  A series of pilot studies were initiated for site selection, gear testing, and survey timing. The culmination of these efforts from 2014-2018 was the standardized nearshore Coastwide Rod-and-Reel Survey which was initiated in 2019.  Since 2019, the WDFW has conducted an annual spring rod-and-reel survey targeting semi-pelagic bottomfish at 125 fixed stations (e.g., index stations) off the Washington coast.  At each station, 5 anglers deploy standardized fishing rigs consisting of 2 shrimp flies and drift over the rocky habitat 3 times.  Each drift is approximately 8 minutes long.  Details regarding sample frame, site selection, and survey methodology for this survey can be found in the Groundfish Subcommittee of the Science and Statistical Committee Visual-Hydroacoustic Survey Methodology Review and Hook-and-Line Survey Workshop details from September 2022 (Reference document??). 

For this assessment abundance indices using data from the two sampling programs described above were evaluated in two ways. First, data from the two projects were evaluated independently. The Black Rockfish Tagging Program tag release data were filtered for sampling events from 1998-2018 in MCA 2 during the spring months (March-July). This time series had the most consistent survey objectives and sampling protocols during the Program. Because black rockfish were explicitly targeted during these trips, no other filters were applied. Catch of black rockfish per angler hour was the response variable, which was an improvement from the past assessment, with covariates year and MCA. CPUE data analysis was done using a hurdle negative binomial regression model.  Figure XX. shows the abundance estimates and 95% confidence intervals for each year. The standardized nearshore Coastwide Rod-and-Reel Survey data were also analyzed using a hurdle negative binomial regression model for data from years 2019, 2021, and 2022.  Data from 2020 were excluded because only index stations in MCA2 were sampled prior to the survey being cancelled due to the Covid-19 pandemic.  Covariates for these analyses included year, MCA, and depth. Figure XX. shows the abundance estimates and 95% confidence intervals for each year.

Secondly, data from the two projects were evaluated concurrently.  Because sampling for both projects targeted black rockfish using rod-and-reel methods on rocky habitats, CPUE data for black rockfish was combined.  Data were filtered for sampling only in MCA 2 during spring months.  Additionally, because of changes in survey designs from 2014-2018, only sets that were within 1km of any 1998-2013 central coast Tagging Program set in MCA 2 were included.  The 1km buffer eliminated most sets done on sand or areas that would not have been fished in a tagging objective set. The index calculation did not include a depth covariate as depth was not recorded during the Tagging Program sampling.  Data were analyzed using a hurdle negative binomial regression model for years 1998-2022 (no sampling occurred in MCA 2 in 2008 or 2017). Figure XX. Shows the abundance estimates and 95% confidence intervals for each year.

Model runs investigating the sensitivity to independent and combined indices for the fishery-independent research sampling projects were considered. Results of the sensitivity runs found……

#### Olympic Coast National Marine Sanctuary adult and young of the year (YOY) surveys

Two surveys conducted between 2015 and 2022 from waters within the Olympic Coast National Marine Sanctuary (OCNMS) were provided for the first time. The adult survey uses SCUBA and belt transects to estimate black rockfish abundance, with fish <10cm considered. Detailed description of survey methods and aims are found in Tolimieri et al. [-@tolimieri_changes_2023] and in a short description (contained in the supplemental materials on this assessment) provided by Ole Shelton (NWFSC), who kindly provided this data for consideration. The adult survey also supplies coarsely binned () length compositions that are used to estimate survey selectivity. The YOY survey is interpreted as an index of recruitment, though admittedly a rough one as it combines yellowtail and blac rockfishes because they are indistinguishable visually at that size and age. For the purposes of this assessment, these data are included but not expected to provide strong signals. It is more to see if the trends in these data are consistent with the trends in the overall assessment. 

<!--chapter:end:21s-.Rmd-->

### \acrlong{s-aslope}

The \gls{s-aslope} operated during the months of October to November aboard the R/V _Miller Freeman_.
Partial survey coverage of the US west coast occurred during the years 1988-1996 and complete coverage (north of 34\textdegree 30\textquotesingle S) during the years 1997 and 1999-2001.
Typically, only these four years that are seen as complete surveys are included in assessments.

<!--chapter:end:21s-aslope.Rmd-->

### \acrlong{s-ccfrp}

Since 2007, the \gls{s-ccfrp} has monitored several areas in California to evaluate the performance of \glspl{mpa}
and understand nearshore fish populations
[@Wendt2009; @Starr2015].
In 2017, the survey expanded beyond the four \Gls{mpa}s in central California
(A&ntilde;o Nuevo, Point Lobos, Point Buchon, and Piedras Blancas)
to include the entire California coast.
Fish are collected by volunteer anglers aboard \glspl{cpfv} guided by one of the following academic institutions based on proximity to fishing location:
Humboldt State University;
Bodega Marine Laboratories;
Moss Landing Marine Laboratories;
Cal Poly San Luis Obispo;
University of California, Santa Barbara; and
Scripps Institution of Oceanography.

Surveys consist of fishing with hook-and-line gear for 30-45 minutes within randomly chosen 500 by 500 m grid cells within and outside \glspl{mpa}.
Prior to 2017, all fish were measured for length and release or descended to depth;
since then, some were sampled for otoliths and fin clips.


<!--chapter:end:21s-ccfrp.Rmd-->

### \acrlong{s-tri}

The \gls{s-tri} was first conducted by the \gls{afsc} in 1977, and the survey continued until 2004 [@weinberg_2001_2002].
Its basic design was a series of equally-spaced east-to-west transects across the continential shelf from which searches for tows in a specific depth range were initiated.
The survey design changed slightly over time.
In general, all of the surveys were conducted in the mid summer through early fall.
The 1977 survey was conducted from early July through late September.
The surveys from 1980 through 1989 were conducted from mid-July to late September.
The 1992 survey was conducted from mid July through early October.
The 1995 survey was conducted from early June through late August.
The 1998 survey was conducted from early June through early August.
Finally, the 2001 and 2004 surveys were conducted from May to July.

Haul depths ranged from 91-457 m during the 1977 survey with no hauls shallower than 91 m.
Due to haul performance issues and truncated sampling with respect to depth, the data from 1977 were omitted from this analysis.
The surveys in 1980, 1983, and 1986 covered the US West Coast south to 36.8\textdegree N latitude and a depth range of 55-366 m.
The surveys in 1989 and 1992 covered the same depth range but extended the southern range to 34.5\textdegree N (near Point Conception).
From 1995 through 2004, the surveys covered the depth range 55-500 m and surveyed south to 34.5\textdegree N.
In 2004, the final year of the \gls{s-tri} series, the \gls{nwfsc} \gls{fram} conducted the survey following similar protocols to earlier years.

<!--chapter:end:21s-tri.RMd-->

### \acrlong{s-wcgbt}

The \Gls{s-wcgbt} is based on a random-grid design;
covering the coastal waters from a depth of 55-1,280 m [@bradburn_2003_2011].
This design generally uses four industry-chartered vessels per year assigned to a roughly equal number of randomly selected grid cells and divided into two 'passes' of the coast.
Two vessels fish from north to south during each pass between late May to early October.
This design therefore incorporates both vessel-to-vessel differences in catchability,
as well as variance associated with selecting a relatively small number (approximately 700) of possible cells from a very large set of possible cells spread from the Mexican to the Canadian borders.

<!--chapter:end:21s-wcgbts.Rmd-->

## Biological Data

The major biological inputs to the models are age and growth parameters, natural mortality, weight-length, maturity and stock-recruitment parameters (Table 40 (OR), and Table 63 (WA)).


### Natural Mortality

Previous assessments of black rockfish used different rates of natural mortality for males versus females to account for the lack of older females in fishery samples (Figure 5 and Figure 6).  The assumed instantaneous rate of natural mortality (M) for males was 0.16 and was constant with age.  Females assumed the male natural mortality value up to age 10, after which natural mortality linearly increased to 0.24 (up to age 15), remaining constant for all subsequent ages. The lower of the two natural mortality rates corresponds to a longevity of only 27 years. Given that individuals over the age of 30 are not uncommon in the age samples being used in the assessment (Figure 3), this seems to be an overestimate of what M should be for black rockfish (Figure 5). Furthermore, the major increase in female natural mortality is invoked well before the theoretical age of asymptotic length, a theory which seems very unlikely given what is known about rockfish life history and growth. This decline in female sex ratio is also seen in other rockfishes (e.g., canary and yellowtail rockfish). The most recent canary rockfish assessment also invoked a similar ramp up to natural mortality to explain this discrepancy on the data (Figure 5).

The pre-STAR approach in this assessment was to fix values of natural mortality at ages to be more in line with the proposed longevity of black rockfish (56 years old; [@love_2011]) and coupled with an age-based selectivity for females to account for the missing samples of that sex (see Section 2.3for details).  A variety of natural mortality estimates based on longevity values and von Bertalanffy values were explored (Table 3).  The estimators using the von Bertalanffy values were ultimately chosen because they incorporated both sex and area differences in natural mortality. This choice of dome-shaped age-based selectivity and fixed lower natural mortality causes the productivity of the population to drop greatly relative to assuming a logistic selectivity and a ramp in M, but also creates a very large amount of cryptic (i.e. unavailable) biomass (Figure 7) compared to the model with a ramp in M (Figure 8). Given such high relative cryptic biomass when invoking dome-shaped age-based selectivity, but discomfort with the M ramp scenario, it was decided during the STAR panel that estimating a sex-specific M was a defensible approach in lieu of data supporting either dome-shaped selectivity and high cryptic biomass or extreme natural mortality values using a ramp in M, and thus was adopted as the approach to M in the California and Washington base models. Sensitivities to the other proposed treatments of natural mortality were also provided.


### Growth (Length-at-Age)

Age and length data were available for all states across years, so an initial investigation of the growth parameters across areas was produced to look at spatial and temporal trends by sex. The standard von Bertalanffy growth function was used to fit the length and age data. Washington State had the most years of available composition data (Figure 2). In general, males seem to be consistently better sampled in all states (Figure 3). Patterns of growth between sexes are similar among areas, while trends in parameters estimates are apparent in California and Washington (Figure 25 and Figure 193).  Washington was the best temporally and numerically sampled area, with California having the fewest samples. Given the level of data, attempts were made to estimate growth parameters internal to the model. When that was not possible, the parameters were fixed to the external fits (Table 18 (CA); Table 40 (OR); Table 63 (WA)).


### Ageing Precision and Bias

Ageing otoliths, while a common practice, rarely provides a prefect estimate of true age. This ageing error (both in bias and imprecision) can have large effects on stock assessment outputs and should be incorporated when using ages. Several multiple age read studies were available to develop ageing error vectors for use in interpreting conditional ages at length. For Washington, there were two data sets: 1) 280 triple reads from WDFW for the commercial fisheries and 2) 3240 double reads from WDFW for the recreational fishery. For Oregon, the Cooperative Ageing Project (CAP) provided a set of 302 multiple reads (five total readers), while ODFW provided 150 from five readers. California had one set of 318 double reads from the Cooperative Ageing Project (CAP).). Resultant forms for each chosen model are given in Figure 4.

The Punt et al. [-@punt_quantifying_2008] method and accompanying software was used to determine the underlying true age distribution and resultant imprecision, assuming at least one of the readers is unbiased. The first reader in all comparisons was assumed unbiased, but we considered several model configurations based on the functional form (unbiased, linear or curvilinear) of bias in the subsequent readers, and precisions of all readers (constant CV, curvilinear standard deviation, or curvilinear CV). Model selection was based on Akaike Information Criterion (AIC) corrected for small sample size (AICc), which converges to AIC when sample sizes are large. Both Washington data sets supported a linear bias in the other readers and constant CV of precision for all readers (Table 2). Oregon data sets agreed that linear bias and curvilinear precision was best supported (Table 2). The California data set supports curvilinear bias in the second reader with constant CV for both readers (Table 2). Resultant forms for each chosen model are given in Figure 4.


### Length-Weight Relationship

Length-weight relationships (kg to fork lengths) were developed with state-specific data. The California weight-length relationship was based on 8943 combined sex samples (Table 18). Oregon relationships were based on length and weight measurements from almost 4,000 individual black rockfish of combined sex and was the same as used in the previous assessment (Table 40). Washington relationships were sex-specific and based on 1551 female samples and 1284 males samples (Table 63), though both sexes had very similar relationships.


### Maturation and Fecundity

The black rockfish maturity is assumed to be based on length, as assumed in past assessments. A notable difference in our approach in these assessments is the use of functional maturity instead of sexual maturity (the typical application). Functional maturity is a more stringent definition of maturity as compared to sexual maturity. Instead of just using the presence of yolk as a measure of sexual maturity, functional maturity takes into account both the presence of strict spawning individuals and the level of atresia or skipped spawning. Such an approach yields length at 50% maturity estimates larger than the standard sexual maturity application. Melissa Head of the Northwest Fisheries Science Center provided estimates of both functional and sexual maturity from black rockfish sampled in the months of July to January (deemed the best time for identifying mature individuals) off Oregon and Washington waters combined. The logistic fit to those the functional and sexual maturity values are found in Table 4. The functional maturity values were sampled from individuals from one year, which happened to be the year of the "warm blob". It is therefore unknown how this may have affected the estimate of functional maturity, particularly the levels of mass atresia in older individuals. Discussion in the STAR panel lead to the decision to assume all individuals above 45cm to be mature in the functional maturity data set in order to be conservative on the estimate of the effect of atresia on maturity estimation. The functional maturity values from this modified data set are used for all states, but sensitivities to the new sexual maturity estimate, as well as the respective area values from the previous assessments and the unmodified functional maturity data set are explored.

Similarly, this assessment, like previous assessments, assumes that weight-specific fecundity is linearly related to female body weight. Values for the slope and intercept were taken from Dick [-@dick_modeling_2009] and are found in the parameterization tables for each state assessment model (Table 18 (CA); Table 40 (OR); Table 63 (WA)). 


### Stock-recruitment function and compensation

The Beverton-Holt stock-recruit model [@beverton_holt_1957] has been the traditional recruitment function for rockfishes and is assumed for black rockfish.  Specifically, the re-parameterized Beverton-Holt that uses a steepness parameter defined as the proportion of average recruitment for an unfished population expected for a population at 20% of unfished spawning output (Mace and Doonan) was used in these assessments.  This is a notoriously difficult parameter to estimate, thus several attempts to derive a prior of steepness have been attempted [@myers_etal_1995; @dorn_advice_2002].  This prior is typically updated each cycle (following the method of Dorn 2002) and subject to a review by the Council's Science and Statistical Committee.  The prior for 2015 has an expected value of 0.773 and a standard deviation of 0.147 using a beta distribution. Attempts were made to estimate this value, but were not successful, so it is fixed and its influence is explored via a likelihood profile.


### Sex Ratio

No information on the sex ratio at birth was available so it was assumed to be 50:50. 

<!--chapter:end:22biology.Rmd-->

## Environmental and Ecosystem Data

<!--chapter:end:23enviro.Rmd-->

# Assessment Model

## Summary of Previous Assessments and Reviews

### History of Modeling Approaches Used for this Stock

#### Black Rockfish South of Cape Falcon

The first stock assessment of black rockfish off Oregon [@stewart_status_1993], which was limited in geographic scope to the northern portion of Oregon, was a Cohort Analysis based on age composition data collected from fish landed at Garibaldi.  The first comprehensive analysis of the black rockfish stock off Oregon and California was by Ralston and Dick [-@ralston_status_2003], who developed a statistical catch-at-age model using Stock Synthesis.  Sampson (2007) used a similar model configuration and approach.

In the 2007 assessment model the data were organized into three basic gear-types (Hook-and-Line, Trawl, and Recreational), the data from Oregon and California were kept separate, and the tuning indices were recreational angler CPUE series based on the same or similar data sources (MRFSS for both states, ORBS for Oregon, and CPFV surveys for California).  Fishing effort was measured in terms of angler-days rather than the angler-hours metric used in the current California and Oregon regional assessment models.  The 2007 assessment used the ODFW tagging study estimates of black rockfish abundance off Newport as a relative abundance index.  Those data were unavailable for the 2003 assessment.  The 2007 assessment also used a juvenile rockfish pre-recruit index, which was unavailable for the previous assessment.

The landings data series in the 2007 assessment differed quite substantially from the series developed by Ralston and Dick for the 2003 assessment.  Neither of those assessments attempted to account for discards, instead assuming that discards were negligible.


#### Black Rockfish North of Cape Falcon

Three full assessments for black rockfish, conducted in 1994, 1999, and 2007, modeled the black rockfish population found in coastal waters between Cape Falcon, Oregon and north to the U.S./Canadian border [@wallace_status_1994; @wallace_status_1999; @wallace_status_1999; @wallace_status_2007].  There have been no update assessments for black rockfish resources.  

The 1994 assessment utilized a Stock Synthesis model configuration, with two auxiliary data sets as black rockfish abundance indicators, one based on tagging CPUE and one on based coastal recreational bottomfish directed effort [@wallace_status_1994].

Wallace et al [-@wallace_status_1999] constructed an assessment model by using the AD Model Builder software (ADMB; [@fournier_1997]) to assess black rockfish abundance.  Three key features of the 1999 model were (1) the parameterization of the expected catches at age, (2) the definitions of the sampling units for the different types of data inputs, and (3) the integration of tagging data explicitly.  The parameterization chosen mostly affected parameter bias whereas the sampling unit designation mostly affected estimator variance.  Both bias and variance were components of overall parameter uncertainty.  The parameterization and the sampling unit definitions were both designed to conform to the actual sampling protocol used, thereby propagating sampling uncertainty through to the final biomass estimates.

The 2007 assessment [@wallace_status_2007] employed Stock Synthesis 2.  Unlike the 1999 assessment, CPUE from the tag release trips and Petersen tagging study abundance estimates were included as relative abundance indices.

#### California, Oregon, and Washington Assessments

The 2015 assessment defined three distinct stocks for assessment. Each stock matched the state boundaries of California, Oregon and Washington. All assessments used the Stock Synthesis 3 version 3.24V. The Washington model had three fisheries (two commercial and one recreational) and considered two surveys. There were three primary data likelihood components for survey indices, lengths and ages. Fits to catches also contribute to the total likelihood, but is typically very small. The models was tuned using the Francis [-@francis_data_2011] method for biological compositions and added variance for survey indices. Recruitment deviations were estimated. Natural mortality was treated as constant and sex-specific, with females having a higher natural mortality than males. 

Results for the Washington assessment of black rockfish in 2015 estimated stock status in that year was 43%, and never showed a decline below the target biomass.


### Most Recent STAR Panel and SSC Recommendations

The STAR panel identified the following issues as sources of major uncertainty:

* Natural mortality, especially in females. There is no data to differentiate whether the missing older females are dying or are avoiding capture. The choice between using a constant (as used in the California and Washington assessments) or step function (used in Oregon) is also an point of uncertainty. 
* The level of cryptic biomass. This is a result of using dome-shaped selectivity to explain the absence of old females.
* Uncertainty in historical catch, especially in the historical trawl fishery.
* Acknowledging that there remains uncertainty in the stock-recruit relationship parameters (particularly the Beverton-Holt steepness parameter).

Most of the above recommendations were included in the 2015 assessment research recommendations. Additionally, stock structure for black rockfish was highlighted as a topic for further consideration, as was the development of a nearshore fishery-independent survey.


# Model description


## Modelling Platform
Stock Synthesis version 3.30.16 was used as the statistical catch-at-age modelling framework. This framework allows the integration of a variety of data types and model specifications. The SS-DL tool (https://github.com/shcaba/SS-DL-tool) was used for model exploration, likelihood profiling, and sensitivity analyses. The companion R package r4ss (version 1.38.0) along with R version 4.0.5 were used to investigate and plot model fits. 


## Bridging the assessment model from Stock Synthesis 3.24 to 3.30 

Since several years have passed from the last assessment model, the Stock Synthesis (SS) modelling framework has undergone many changes. While the specific changes in the model can be found in the model [change log](https://github.com/nmfs-stock-synthesis/stock-synthesis/blob/v3.30.19/Change_log_for_SS_3.30.xlsx?raw=true), here we simply update the model from the older 3.24V version to the newer 3.30.20 version. The point here is to present any differences in the model outputs when using the same information. This was first done by migrating the data and parameter specifications from the former files to the newer files. This migration was assisted using the [SS-DL tool](https://github.com/shcaba/SS-DL-tool). Once the old data was transferred to the SS 3.30.20 file, two versions of the model were ran. 

  1) Fixing all parameter values to the values found in the 2015 model. 
  2) Allowing the same parameters estimation specification as in the 2015 model. 

Results are similar between models when all parameters are fixed from the 2015 model in the updated SS files, although there are scale differences (Figure \ref{fig:ssb_bridge_comps}) and small relative stock status differences (Figure \ref{fig:deps_bridge_comps}) when the new SS version is allowed to estimate the same parameters as estimated in the 2015 version. These model comparisons are adequate to move ahead using the newest version of SS 3.30.20 without expecting large differences in reference models being due to versions of SS. 




<!--chapter:end:30model.Rmd-->


<!--chapter:end:31summary.Rmd-->

## Model Structure and Assumptions


### Model Changes from the Last Assessment (not required for an update assessment)


### Modeling Platform and Structure

### Fleet and survey designations

The Oregon model is structured to track several fleets and include data from several surveys:

* Fleet 1: Trawl commercial fishery
* Fleet 2: Non-Trawl live fish commercial fishery
* Fleet 3: Non-Trawl commercial fishery: mainly hook-and line fishery
* Fleet 4: Recreational ocean fishery
* Fleet 5: Recreational shore fishery
* Survey 1: Onboard CPFV CPUE survey
* Survey 2: Tagging abundance survey
* Survey 3: MRFSS CPUE survey
* Survey 4: ORBS CPUE survey
* Survey 5: Commercial logbook CPUE survey
* Survey 6: Research survey: small fish 

The Washington model is structured to track several fleets and include data from several surveys:

* Fleet 1: Trawl commercial fishery
* Fleet 2: Non-Trawl commercial fishery: mainly hook-and line fishery
* Fleet 3: Recreational fishery
* Survey 1: Dockside CPUE survey
* Survey 2: Tagging CPUE survey

### Model likelihood components

There are four primary likelihood components for each assessment model:
1. Fit to survey indices of abundance.
2. Fit to length composition samples.
3. Fit to age composition samples (all fit as conditional age-at-length).  
4. Penalties on recruitment deviations (specified differently for each model).

Indices of abundance are assumed to have lognormal measurement errors. Additional variance to the inputted log-standard deviation is estimated in the base case (Washington).  Length compositions and conditional age at length samples are all assumed to follow a multinomial sampling distribution, where the sample size is fixed at the input sample size calculated during compositional example, and where this input sample size is subsequently reweighted to account for additional sources of overdispersion (see Section 2.3.3 below).  Recruitment deviations are assumed to follow a lognormal distribution, where the standard deviation of this distribution is tuned as explained below.


### Data Weighting

The reference model estimates additional variance on the Oregon recreational survey data to allow the model to balance model fit to that data while acknowledging that variances may be underestimated in the index standardization. The input CVs range from 1%-7%, which is very small (Table \ref{tab:OR_vermilion_ORBSindex}). A sensitivity was run with no extra variance estimated, as well as removal of the index data.

Initial sample sizes for the commercial and recreational length and conditional age-at-length compositions were based on the number of fish sampled. The method of Francis [-@francis_data_2011, equation TA1.8] was then used to balance the length and conditional age-at-length composition data among other inputs and likelihood components. The Francis method treats mean length and age as indices, with effective sample size defining the variance around the mean. If the variability around the mean does not encompass model predictions, the data should be down-weighted until predictions fit within the intervals. This method accounts for correlation in the data (i.e., the multinomial distribution), but can be sensitive to years that are outliers, as the amount of down-weighting is applied to all years within a data source, and are not year-specific. Sensitivities were performed examining different data-weighting treatments: 1) the Dirichlet-Multinomial approach [@thorson_model-based_2017], 2) the McAllister-Ianelli Harmonic Mean approach [@mcallister_bayesian_1997], or 3) no data-weighting of lengths.



### Model Parameters


### Key Assumptions and Structural Choices

All life history parameters are estimated except the CV at length at $t_0$. Estimated parameters in the model are natural mortality ($M$) and all growth parameters ($L_{\infty}$, $k$, $t_0$, CV at $L_{\infty}$; length CV at $t_0$ was fixed as it had little impact on the model) were estimated, as were the two selectivity parameters for each fleet and the survey, the log of the initial recruitment ($logR_0$), and recruitment deviations. Length at maturity, fecundity-weight, and length-weight relationshop, steepness ($h$) and recruitment variance were all fixed. Sensitivity scenarios and likelihood profiles were used to explore uncertainty in the values of the natural mortality and growth parameters. When estimating parameters, the prior for natural mortality was assumed lognormal with a standard deviation of 0.438 (based on the prior developed using the Natural Mortality Tool (see Biology section for more details); growth parameters were estimated with no priors.

OREGON: While the step in M is similar in concept to the past models use of a ramp in M, the magnitude of that step is much smaller (step from 0.17 to 0.20, rather than a ramp from 0.16 to 0.24). Selectivity also differs from the last model, as well as from the California and Washington models, in the use of both sex-specific length- and age-based selectivity forms. Selectivity for the ascending portion of the selection curves for all five of the fleets was modeled using length-based selection with no differences in length selection by sex.  The trawl fishery assumed asymptotic for both sexes, but with the allowance of a female offset to male selectivity. The live fish fishery selectivity was shared by both sexes (dome-shaped), but the dead fish fishery was modeled as a female offset in the dome-shaped parameters. The recreational ocean fishery used an age-based selectivity offset on the descending limb for females relative to males, which were assumed fully-selected at all ages (length-selectivity was used to describe the active male selectivity in this fishery). Similar to the live fish fishery, the recreational shore was dome-shaped and shared for both sexes. Finally, the catchability parameter for the tagging study was fixed to 0.25.

WASHINGTON: The most dramatic model specification in these models, in relation to past assessments, is the choice to estimate sex-specific natural mortality rather than assuming dramatic changes (i.e. a ramp) in natural mortality. This adjustment has major implications on stock productivity and necessitates the exploration of alternative runs that assume the former treatment of ramping natural mortality for females as well as the possibility that female cryptic biomass exists via age-based dome-shaped selectivity. The performance of each state model with the removal of each data type is also provided to give support for the final choice of the base models, which try to balance the realism (i.e., do the results makes sense?) with parsimony (i.e., are we trying to do too much with the model?).


## Model Selection and Evaluation
The base assessment model for Oregon `r spp` was developed to balance parsimony and realism, and the goal was to estimate a spawning output trajectory and realtive stock status for the population of `r spp` in state and federal waters off Oregon. The model contains many assumptions to achieve parsimony and uses different data types and sources to estimate reality. A series of investigative model runs were done to achieve the final base model. These include considerations of model structure, data and parameter treatment, estimation phasing, and jittered starting values to achieve a converged and balanced model that provides sensible parameter estimates and derived quantities.


<!--chapter:end:32structure.Rmd-->

## Model Diagnostics


### Model Convergence and Acceptability

While there is no definitive measure of model convergence, several measures are routinely applied. These criteria include a low maximum gradient (`r model$maximum_gradient_component`), inversion of the Hessian (passed), reasonable parameter values (passed), and acceptable fits to data (passed). 

An extra effort was given to ensure the model did not rest on a local likelihood minimum. This was done by starting the minimization process from dispersed parameter values away from the maximum likelihood estimates to determine if the approach found a better model fit (i.e., minimum negative log-likelihood value). Starting parameters used a jitter shift value of 0.001. This was repeated 100 times with 12 out of 100 runs returned to the reference model likelihood (Figure \ref{fig:jitter_001}). A better fit, lower negative log-likelihood model was not found in any of remaining 96 runs. The model did not experience convergence issues when provided reasonable starting values. Through the jittering and likelihood profiles, the present reference model represents the best fit to the data given the assumptions made. 

Models were considered converged if they meet low gradient requirements (<0.001) and produced asymptotic standard deviations. Additional explorations for a consistent likelihood minimum was performed using jittered (0.1) starting values. A total of 100 jittered runs were performed for each model. Across all jittered runs, the lowest likelihoods of each respective model matched the base case likelihood (Figure 99 (OR); Figure 195 (WA)).


## Base Model Results


### Parameter Estimates

The final base model incorporated aspects of both models proposed to the Mop-up STAR Panel. Below is a list of the more salient specifications (See also Section 2.4.1.2 for comparisons with the California and Washington models, of which the Oregon model shares many similar specifications when not listed below):
* Female M fixed at 0.17, with a step at age 10 to 0.2.
* Male M fixed at 0.17 for all ages.
* Tag Q fixed at 0.25.
* Selectivity for the trawl fishery was logistic and length-based, but allowed to be sex-specific. 
* Selectivity for the ocean-boat recreational fishery was estimated as age-based with dome-shaped parameters for females.
* Conditional age-at-length was weighted using harmonic means.
* Recruitment deviations were not estimated.	


### Fits to the Data

The Oregon base case model shows the indices trending upward in the last three years, with the exception of the Commercial Logbook index.  Fits to the indices were generally poor, with mostly straight-line fits to most of them (Figure 100 to Figure 104).  

Length composition fits are good for the live and dead non-trawl fisheries and for the recreational ocean-boat fishery, which represent the bulk of the data, but poor for the trawl and small fish research study, both of which had small samples sizes (Figure 105  to Figure 119).  Francis weighting are shown in Figure 120 to Figure 124. Fits to the weighted conditional age-at-length compositions show generally good agreement between observed and expected values (Figure 125 to Figure 132).  Estimated selectivity curves show that the trawl fishery is asymptotic in length for males and females, dome-shaped for both sexes in the live and dead fish fishery and the small fish for research survey, and dome-shaped and age-based for females in the recreational ocean fishery (Figure 147 to Figure 149).

Sex-specific growth was estimated in the base model (Figure 150, Table 40).  Females grew bigger than males (Figure 149) and both were similar to the values estimated when the data was fit outside the model.

A total of sixteen primary parameters were estimated, along with sixty recruitment deviations. The reference model parameter estimates along with asymptotic standard errors are shown in Table \ref{tab:model-param} and the likelihood components are shown in Table \ref{tab:likes}. Estimates of derived reference points and approximate 95 percent asymptotic confidence intervals are provided in Table \ref{tab:referenceES}. 

The natural mortality for females and males was estimated at `r round(model$parameters[model$parameters$Label == "NatM_p_1_Fem_GP_1", 3],3)` and `r round(model$parameters[model$parameters$Label == "NatM_p_1_Mal_GP_1", 3],3)` yr^-1^, respectively. These values are below the mean prior value, but not unreasonable given the corresponding longevities would be between 67 and 75 years old and the sampled maximum age of 68 came from a fished population.

The estimates of sex-specific growth parameters showed some differences from the externally estimated starting values (Table \ref{tab:model-param} and Figure \ref{fig:len-age-ss}). While $L_{\infty}$ was similar to the external estimates, the model estimated $k$ for female and male fish were greater than the values estimated externally (`r round(model$parameters[model$parameters$Label == "VonBert_K_Fem_GP_1", 3],3)`  for females and `r round(model$parameters[model$parameters$Label == "VonBert_K_Fem_GP_1", 3]*exp(model$parameters[model$parameters$Label == "VonBert_K_Mal_GP_1", 3]),3)` yr^-1^ for males).  The majority of female and male `r spp` growth occurs at younger ages, reaching near maximum length by age 20-25, depending upon sex, with female `r spp` reaching larger maximum lengths (Figure \ref{fig:len-age-ss}). 

The estimated logistic selectivity curves for the commericial and recreational fishery look plausible (i.e., as a model convergence check for realism, the selectivity curves are not overtly outrageous) for each fishery and are very similar to each other (Figure \ref{fig:fleet_selectivity}). Length at 50% selectivity (commercial = `r round(model$parameters[model$parameters$Label == "Size_DblN_peak_Commercial(1)", 3],2)` cm; recreational = `r round(model$parameters[model$parameters$Label == "Size_DblN_peak_Recreational(2)", 3],2)` cm) was between the length at 50% (39.4 cm) and 95% maturity (48 cm). Future assessments could opt for parsimony and combine these two fisheries into one combined selectivity, though the model had no issue adding two more parameters given the available length data.

The time series of estimated recruitments and annual recruitment deviations are shown in Figures \ref{fig:recruits} and \ref{fig:rec-devs}. Years with the highest recruitment deviations were estimated to have occurred in 1993-1994, 1998, 2005, and 2015. The variance check on the recruitment deviations indicates well informed recruitments from 1980 to 2015, providing justifation for the estimation of recruitment. Recruitment deviations after 2015 are relatively uninformed with estimated deviations near zero where recruitment is estimated primarily based on the spawner-recruit curve (Figure \ref{fig:bh-curve}). The recruitment bias adjustment applied within the model across years is shown in Figure \ref{fig:bias-adj}.


### Population Trajectory

The Oregon base case model shows the indices trending upward in the last three years, with the exception of the Commercial Logbook index.  Fits to the indices were generally poor, with mostly straight-line fits to most of them (Figure 100 to Figure 104).  

The predicted spawning output (in millions of eggs) is provided in Table \ref{tab:timeseries} and plotted in Figure \ref{fig:ssb}. Estimated spawning output shows a large decline starting in the 1970s, with a continued decline into the late 1990s. This tracks the time period of major removals, though removals have stayed somewhat elevated since. Strong recruitments since the 1990s have supported the elevated catches. The estimate of total biomass over time, which tracks that of spawning output, is shown in Figure \ref{fig:tot-bio}.

Relative spawning output declined below the management target ($SB_{40\%}$) by the 1990s, but quickly rebounded to high relative spawning output, but has declined over the past 10 years (Figure \ref{fig:depl}). The relative stock status at the start of `r endyr + 1` (`r round(model$derived_quants[model$derived_quants$Label == 'Bratio_2021','Value'], 2)`) is estimated to be well above the rockfish relative biomass target of 0.4. Uncertainty intervals are wide given the number of estimated parameters, and indicate the population never goes below the management target ($SB_{40\%}$). The strong recruitment events that are supported by all fishery-dependent data sources are responsible for the dramatic increase and elevated stock status. Numbers of age-0 individuals indcate those years of particularly strong recruitment (Figure \ref{fig:recruits}). 


WA:
The Washington base case model showed acceptable and dynamic fits to both indices (Figure 196 and Figure 197). The dockside CPUE was the most informative and best fit of the two series.  Fits to the length compositions are generally good (Figure 198 to Figure 210). The worst fits are found in data with low effective weight, such as the trawl samples. The Francis weighting shows a good match between expected and observed mean lengths over time for all fleets (Figure 207to Figure 210). Fits to the unweighted conditional age-at-length compositions shows generally good agreement between observed and expected ages at length (Figure 211 to Figure 216). Selectivity curves fits indicate the trawl fishery to be very different than other fleets (Figure 217). The trawl fishery effectively sampled a smaller portion of the adult population than the other fleets (Figure 219). Sex-specific growth was estimated in the base model (Figure 220). Natural mortality estimates were much larger than the initial prior value (Table 63). Numbers at age table and plots are given in the Supplementary tables (worksheets "WA #s at Age" and "WA #s at age plots").

<!--chapter:end:33results.Rmd-->


### Sensitivity Analyses

Several sensitivity runs were considered for the Oregon model. The first group of sensitivity runs (scenarios 1-6; Table 43; Supplemental table "OR Sensitivities- Like Comps") sequentially removes indices. The second group (scenarios 7-13; Table 43; Supplemental table "OR Sensitivities- Like Comps") sequentially removes length composition data. The third group (scenarios 14-19; Table 43; Supplemental table "OR Sensitivities- Like Comps") sequentially removes indices, size or age composition data. The last group (scenarios 19-33; Table 44; Supplemental table "OR Sensitivities- Model specs") covers a variety of issues: 

1. Estimate M step
2. M step (at age 10) fixed at 2007 values
3. M ramp (at age 10) fixed at 2007 values
4. M estimated, no step
5. M using Then et al. 2015 VBGF calculation; age-based selectivity
6. M based on Then et al. (2015) using amax = 56; age-based selectivity
7. M based on Hamel (2015) using amax = 56; age-based selectivity
8. Estimate female M with male offset set at -0.2
9. Fix tag Q = 12.5%
10. Estimate tag Q
11. Data weighting: length using Francis method; no age weighting
12. Data weighting: lengths and ages using harmonic mean
13. No recruitment estimation
14. No extra variance estimated on all indices 
15. Recreational ocean fishery selectivity logistic

The model was generally insensitive to the removal of indices, except when the tag index was removed as well as the tag Q fixed value. The model was more sensitive to the removal of age compositions, particularly the recreational ocean fishery, resulting mostly in the changes of absolute spawning output. This general insensitivity is due to the fact that tag Q is fixed in the base model. Previous model exploration that did not fix tag Q showed sensitivity to both the length compositions, which favored higher M values, and age compositions, which favored lower M values.

Results of the model specification sensitivity runs demonstrate a strong sensitivity to the scale parameter Q from the tagging data, though selectivity can also change results (Table 44; Figure 153 and Figure 154). The model is very rigid given the fixed value of tag Q. There is very little effect on stock status when changing the parameterization of M, selectivity, data weighting, or the estimation of recruitments (Figure 155 to Figure 158). Only the treatment of the tag Q parameter causes notable results. Because the Oregon model both kills individuals through elevated M values, but also hides the females via dome-shaped selectivity, the biomass unavailable to fishing mortality ("cryptic biomass") was considered for different values of tag Q. Despite very different assumptions of stock scale and status from the different Q parameters (Figure 153 and Figure 154), the cryptic biomass was comparable (Figure 159 to Figure 161). 

Overall, the scale of the population (and thus the resultant OFL at the proxy FMSY value) is the most sensitive derived quantity. Lower natural mortality rates caused big changes in the spawning output. Such lower productivity also caused reduced OFLs. Estimating recruitment and assuming logistic selectivity for the recreational ocean fishery also caused changes in the population scale.

WA:
Several sensitivity runs were considered for the Washington model. The first group of sensitivity runs (scenarios 2-4; Table 65; Supplemental table "WA Sensitivities- Like Comps") sequentially removes indices. The second group (scenarios 5-9; Table 65; Supplemental table "WA Sensitivities- Like Comps") sequentially removes length composition data. The third group (scenarios 10-13; Table 65; Supplemental table "WA Sensitivities- Like Comps") sequentially removes age composition data. The last group (scenarios 14-36: Table 66; Supplemental table "WA Sensitivities- Model specs") covers a variety of issues

1. Estimate growth deviations (1980-2014)
2. Estimate growth blocks (1980-1999; 2000-2014)
3. M ramp in females as in 2007 assessment.
4. Estimate M ramp
5. Fix M to Hamel approach value (0.0964)
6. M from STAR presented base case (based on Then et al. 2015); age-based selectivity
7. M based on Then et al. (2015) using amax = 56; age-based selectivity
8. M based on Hamel (2015) using amax = 56; age-based selectivity
9. Estimate M with dome-shaped, age-based selectivity
10. Estimate M ramp with dome-shaped, age-based selectivity
11. Sexual maturity used in the last assessment (2007)
12. M ramp in females and sexual maturity in 2007
13. Sexual maturity estimated from recent samples (2015)
14. M ramp in females and sexual maturity in 2015
15. Functional maturity using original data set
16. Fecundity is linear with intercept = 0, slope =1.
17. Fecundity from 2007 assessment
18. No recruitment estimation
19. Recruitment estimated all years; ?R tuned
20. Use harmonic mean when tuning length compositions
21. Use harmonic mean when tuning length and conditional age-at-length compositions
22. No extra variance estimated on all indices 
23. Recreational fleet length selectivity is dome-shaped 

Results of the likelihood component sensitivity runs are found in Table 65. The model was most sensitive to the exclusion of the dockside recreational index. It was also sensitive to the removal of all length and age data, and particularly to the recreational age data. 

Results of the model specification sensitivity runs are found in Table 66. The largest sensitivities were found in the treatment of maturity, selectivity and natural mortality. The use of sexual maturity (scenarios 24-27) changed the terminal year scale and status of the stock significantly, making it much less reduced in status, though having a smaller influence on catch at SPR50% (Table 21). Natural mortality scenarios with lower M but age-based selectivity had the biggest effect in spawning output (Figure 225), whereas scenarios with ramping of M caused the biggest changes in (i.e., improving) stock status (Figure 226). Fixing M to low values but not compensating with dome-shaped, age-based selectivity caused the population to crash. Scenarios with ramps in M or M estimated, regardless of selectivity form for females, caused the biggest increases in catch at SPR50%. Recruitment deviations are relatively insensitive to the different natural mortality specifications (Figure 227). Most natural mortality scenarios bring the fishing intensity below the SPR harvest level (Figure 228). 


### Retrospective Analysis

OR: Retrospective runs of for the last 5 years plus and additional retrospective back to 2006 (-8 years, the last year of data available in the previous assessment) were conducted. The Oregon model demonstrates little retrospective patterns in both spawning stock output (Figure 170) and relative spawning stock output (Figure 171). 

WA:Retrospective runs of for the last 5 years plus and additional retrospective back to 2006 (-8 years, the last year of data available in the previous assessment) were conducted. The Washington model demonstrates little retrospective pattern in both spawning stock output (Figure 235) and relative spawning stock output (Figure 236). Both indicate that in general, as data is removed, the stock gains spawning output and the relative stock status increases.


### Likelihood Profiles

OR:
Profiles for natural mortality, steepness, initial recruitment (lnR0), and the ln(tag Q) are shown in Figure 162 to Figure 169. The data in the model strongly support high M values, though fixing the tag Q has removed the sensitivity of the derived quantities to this M values (Figure 162 and Figure 163). Fixing Q also makes all likelihood components support a larger M, whereas when tag Q is estimated, length compositions support higher M values and age composition support lower M values. Profiles over steepness show higher steepness values are most supported and have little influence on the derived quantities (Figure 164). All likelihood components but the abundance indices support higher steepness (Figure 165). The lnR0 parameter is very well defined (Figure 166) and supported consistently by all likelihood components (Figure 167). Higher ln(tag Q values) are most supported by the data, a value very different from the fixed value used in the base model (Figure 168). Derived quantities are very sensitive to the value of this parameter. The ln(tag Q) best supported is most influenced by the age data, with the length data supporting a lower ln(tag Q).

WA:
Likelihood profiles were conducted for sex-specific natural mortality (where males were a fixed offset from females), for population scale (initial recruitment (lnR0)) and stock productivity (steepness (h)). Natural mortality values of females between 0.15 and 0.18 were hard to differentiate in the likelihood value, and did not show much difference is stock scale or status (Figure 229). Overall, derived quantities were very sensitive to natural mortality. The likelihood components degraded most with lower and higher natural mortality values (Figure 230). Age and length likelihood components were opposed in their information on natural mortality. Initial recruitment was well informed, with stock status sensitive to the value of initial recruitment (Figure 231). Values of lnR0 greater than the base case estimate quickly rise to unfished status. Age and length likelihood components were sensitive to the value of lnR0 and opposed each other in the relationship to ln R0 (Figure 232). Regarding steepness, a freely estimated steepness would move towards the high bound (Figure 233). Values below 0.7 were not supported by the data. Derived quantities were fairly insensitive to the value of steepness. Most likelihood components behaved consistently to steepness values, though age compositions were most sensitive to high steepness values while all other components were sensitive to very low steepness values (Figure 234).


### Unresolved Problems and Major Uncertainties

<!--chapter:end:34diagnostics.Rmd-->

# Management 

## Reference Points
OREGON: Spawning output declined as it tracks quickly rising recreational catches in the late 1970s, then stabilizes (Figure 151) Stock status is above the target reference point (SB40%) at 60% (59%-62% 95% asymptotic intervals; Table 45). Unfished spawning output was measured at 1385 (1212-1557 95% asymptotic intervals; Table 45) and spawning output at the beginning of 2015 was estimated to be 838 (740-936 95% asymptotic intervals). According to this model specification, there is no time in which the stock was below the target or limit reference point, but is currently pointed downward. Fishing intensity has been less than the SPR50% rate the whole time series (Figure 173). The phase plot shows the interaction of fishing intensity and biomass targets (Figure 174). The equilibrium curve is shifted left, as expected from the high fixed steepness, showing a more productive stock than SPR50% would suggest (Figure 175). The target stock size based on the biomass target (SB40%) is 554 (millions of eggs), which gives a catch of 556 mt. Equilibrium yield at the proxy FMSY harvest rate corresponding to SPR50% is 518 mt (Table 45).

WA: 
Spawning output declined over a large extent of the time series, with an increasing and or more stable population prevailing since the late 1980s (Figure 221) Stock status is above the target reference point (40%) at 43% (36%-50% 95% asymptotic intervals; Figure 222). Unfished spawning output was measured at 1356 (1228-1483 95% asymptotic intervals; Table 67) and spawning output at the beginning of 2015 was estimated to be 582 (467-698 95% asymptotic intervals) There seems to be no time in which the stock was below the limit reference point and has only fluctuated above the target reference point. Recruitment has fluctuated regularly over the last 25 years (Figure 223 and Figure 224). Despite being above the target biomass, fishing intensity has been above the SPR50% rate in the last couple of years (Figure 237). The phase plot shows the interaction of fishing intensity and biomass targets (Figure 238). The equilibrium curve is shifted left, as expected from the high fixed steepness, showing a more productive stock than SPR50% would suggest (Figure 239). The target stock size based on the biomass target (SB40%) is 542 (millions of eggs), which gives a catch of 337 mt. Equilibrium yield at the proxy FMSY harvest rate corresponding to SPR50% is 311 mt (Table 67).

From Vermilion:
Reference points were calculated using the estimated fishery selectivity and removals in the most recent year of the model (`r endyr`, Table \ref{tab:referenceES}). Sustainable total yields  were `r round(model$derived_quants[model$derived_quants$Label == 'Dead_Catch_SPR', 'Value'], 2)` mt when using an $SPR_{50\%}$ reference harvest rate. The spawning output equivalent to 40 percent of the unfished spawning output ($SB_{40\%}$) was `r round(model$derived_quants[model$derived_quants$Label == 'SSB_SPR', 'Value'], 2)` meggs. 

The `r endyr + 1` spawning output relative to unfished equilibrium spawning output is above the `r spp` relative biomass target of 40 percent (Figure \ref{fig:depl}). The fishing intensity, $1-SPR$, of recent years was near or above the harvest rate limit ($SPR_{50\%}$) for most of the 1980s and 1990s. Recent years also show near target fishing levels (Table \ref{tab:timeseries} and Figure \ref{fig:1-spr}), highlighting how the sustainability of current fishing levels are very sensitive to incoming recruitment. Table \ref{tab:referenceES} shows the full suite of estimated reference points for the base model and Figure \ref{fig:yield} shows the equilibrium curve based on a steepness value fixed at `r round(model$parameters[model$parameters$Label == "SR_BH_steep","Value"], 2)`.


## Unresolved Problems and Major Uncertainties


## Harvest Projections and Decision Tables

A ten year (2023-2032) projection of the reference model with removals in 2021 and 2022 provided by the Groundfish Management Team for each fleet under the category 1 (sigma=0.5) time-varying buffer using $P^*$ = 0.45 and 40-10 ABC control rule is provided in Table \ref{tab:project}.



## Evaluation of Scientific Uncertainty

The estimated uncertainty in the base model around the `r endyr + 1` spawning output is $\sigma$ = `r round(model$Pstar_sigma,2)` and the uncertainty in the base model around the `r endyr + 1` OFL is $\sigma$ = `r round(model$OFL_sigma,2)`. The estimated model uncertainty was less than the category 1 groundfish data moderate assessment default value of  $\sigma$ = 0.5. 


## Research and Data Needs


<!--chapter:end:40management.Rmd-->

# Acknowledgments
Here are all the mad props!

<!--chapter:end:41acknowledgments.Rmd-->

\clearpage

# References
<!-- If you want to references to appear somewhere before the end, add: -->
<div id="refs"></div>
<!-- where you want it to appear -->
<!-- The following sets the appropriate indentation for the references
  but it cannot be used with bookdown and the make file because it leads
  to a bad pdf.
\noindent
\vspace{-2em}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
 -->

<!--chapter:end:49bibliography.Rmd-->

\clearpage
# Tables


<!--chapter:end:52tables.Rmd-->

\clearpage
# Figures

<!-- ====================================================================== --> 
<!-- ******************************* ES *********************************** --> 
<!-- ====================================================================== --> 


<!-- ====================================================================== --> 
<!-- ******************* Intro section *************************** --> 
<!-- ====================================================================== --> 

<!-- ====================================================================== --> 
<!-- ******************* Data Used in the Model *************************** --> 
<!-- ====================================================================== --> 

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "data_plot.png"), 
caption = "Summary of data sources used in the reference model",
label = 'data-plot')
```

<!-- ====================================================================== --> 
<!-- ******************* Bridge Model ************************************* --> 
<!-- ====================================================================== --> 

```{r, results = 'asis'}
add_figure(
  filein = file.path(bridge_fig_loc, "WA_SB_comp_plot.png"), 
  caption = "Comparison of spawning output for black rockfish in waters off of Washington between Stock Synthesis versions 3.24 and 3.30. Uncertainty envelops are 95% confidence intervals",
  label = 'ssb_bridge_comps')

``` 
```{r, results = 'asis'}
add_figure(
  filein = file.path(bridge_fig_loc, "WA_Dep_comp_plot.png"), 
  caption = "Comparison of spawning output for black rockfish in waters off of Washington between Stock Synthesis versions 3.24 and 3.30. Uncertainty envelops are 95% confidence intervals",
  label = 'deps_bridge_comps')

``` 

<!--chapter:end:53figures.Rmd-->

