```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through xxxx.")
```

# Executive summary{-}

## Stock{-}

This assessment reports the status of `r spp` (*Sebastes melanops*) off Oregon state using data through `r endyr`. `r Spp` are also found in California (their core range) and Washington waters of the U.S. West Coast, and those are treated in separate area-based stock assessments given different mangement considerations and exploitation histories as discussed at the preliminary stock assessment workshop in February 2023 [@PFMC_dataworkshop_2023]. There is substantial biogeographic separation in the populations off Oregon and Washington, thus justifying separation of those populations into different management units and stock assessments. `r spp` are also caught from the waters off British Columbia and Alaska, but there have not been any formal assessments of stock status for those areas. Genetic studies of stock structure indicate fish in Alaska are more differentiated than the along the contiguous west coast of the Unites States, and that genetic diversity varies in a nonsystematic way from California to Oregon [@Hessetal_2023].

## Removals{-}

`r spp` are caught by a wide variety of gear types in Oregon and in recent decades have been a very important target species for recreational charter-boats and private sport anglers (i.e., ocean boat fleet).  In recent years the recreational fishery has accounted for most of the `r spp` catches (Figure \ref{fig:es-catch}).  `r spp` can also be an important component of nearshore commercial fisheries, either as incidental catch by the troll fishery for salmon or as directed catch by jig fisheries for groundfish.  Further, there are nearshore commercial fisheries that catch and sell fish live for the restaurant trade.  There has been almost no trawl-caught landings of `r spp` in recent years (Table \ref{tab:removalsES}), but trawl landings in the 1940s to 1970s were more prominant (Figure \ref{fig:es-catch}).

Detailed reports of commercial landings of `r spp` are generally unavailable prior to 1981, when the Pacific Fishery Information Network (PacFIN) database began.  The catch series prior to 1981 for these assessments were derived by applying available estimates or assumed values for the proportion of `r spp` landings in reported landings of rockfish.  Observer data, which are available since the early 2000s, indicate low levels of discarding of `r spp`, generally less than 2% of total catch. While `r spp` are unlikely to have ever comprised a large percentage of overall rockfish landings due to their low abundance compared to other rockfish species, it seems plausible that they have been more than a trivial component due to their nearshore distribution for many years.  `r spp` were one of only four rockfish species mentioned by scientific name in reports of rockfish landings in Oregon during the 1940s.

Overall, removals of `r spp` remained relatively low (less than 100 mt) until the mid to late 1970s when landings quickly quadrupled with the expansion of the recreational fishery. Since the 1980s, removals have consistently fluctuated between 300 and 600 mt (no major trend), comprising mostly of removals from the ocean boat recreational fleet and the non-trawl commercial fleet (Figure \ref{fig:es-catch}).  


\input{tex_tables/a_Catches_ES.tex}


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the reference model where catches in metric tons by fleet are stacked",
label = 'es-catch')
```

\clearpage

## Data and assessment{-}

The first `r spp` stock assessment along the west coast of the United States that included the majority of Oregon waters was completed in 2007, covering the area south of Cape Falcon, Oregon to north of Point Piedros Blancos, California [@sampson_status_2007]. In 2015, a subsequent assessment was completed that included Oregon waters only as one of three (also Washington and California) seperate assessment areas delineated by state lines [@cope_assessments_2016]. Similarly, this assessment treats Oregon waters as a single assessment area. The previous two assessments used Stock Synthesis software, as does this one (version 3.30.21.00).

This assessment integrates data and information from multiple sources into one modeling framework. The stock assessment model for `r spp` is informed by catch data from two commercial fleets and two recreational fleets, six abundance indices, five sets of length composition data, and 3 sets of conditional age-at-length compositions. It also uses an ageing error matrix to incorporate ageing imprecision and applies fixed parameterizations of weight-at-length, maturity-at-length, fecundity-at-length, the Beverton-Holt stock-recruitment steepness value, and recruitment variability. Life history parameters were sex-specific (i.e., a two-sex model) with natural mortality fixed at external estimates and growth and recruitment parameters estimated.  Additional parameters that were estimated include initial population scale ($lnR_0$), selectivity for each fishery and survey, and extra survey variance. The base model was tuned to account for the weighting of the length and age data and index variances (which was estimated), as well as the specification of the recruitment bias adjustments.  Derived quantities include, among other things, the time series of spawning biomass, age and size structure, and current and projected future stock status. The model covers the years `r startyr` to `r endyr`, with a 12 year forecast beginning in `r endyr + 1`.

Within model uncertainty is explicitly included in this assessment by parameter estimation uncertainty, while among model uncertainty is explored through sensitivity analyses addressing alternative input assumptions such as data treatment and weighting, and model specification sensitivity to the treatment of life history parameters, selectivity, recruitment, and survey catchability. A reference model was selected that best fit the observed data while concomitantly balancing the desire to capture the central tendency across those sources of uncertainty, ensure model realism and tractability, and promote robustness to potential model misspecification.

## Stock biomass and dynamics{-}

Spawning output (in millions of eggs; meggs) instead of spawning biomass is used to report the functionally mature population scale because fecundity is nonlinearly related to body female weight. The estimated spawning output at the beginning of 2023 was `r ssb_curr` meggs (~95 percent asymptotic intervals: `r ssb_curr_low` to `r ssb_curr_hi` meggs, Table \ref{tab:ssbES} and Figure \ref{fig:es-ssb}), which when compared to unfished spawning output (`r ssb0`) meggs gives a relative stock status level of `r round(100*depl_curr,0)` percent (~95 percent asymptotic intervals: `r round(100*depl_curr_low,0)` to `r round(100*depl_curr_hi,0)` percent, Figure \ref{fig:es-depl}).  Overall, spawning output declined with the onset of increasing commercial removals in the 1960s and continued to decline with the increase in recreational catches through the 1990s, even dropping below the target relative stock size from 1993 to 2008, before steadily increasing back above target since that time. The largest of the estimated recruitment pulses occurred in 2008 and was followed by several above average recruitment years in the early 2010s, which contributed to the increase in spawning output through the mid to late 2010s. The minimum relative stock size of `r round(100*as.numeric(depl_min),0)` percent of unfished levels is estimated to have occurred in `r depl_min_yr`. Accordingly, the stock has not been below the minimum stock size threshold (i.e., never overfished based on median estimates). Currently the stock is estimated above the management target of $SO_{40\%}$ in 2023 and is estimated to have remained above the target since 2009 (Table \ref{tab:ssbES} and Figure \ref{fig:es-depl}).

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots",
"ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
label = 'es-ssb')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts9_Relative_spawning_output_intervals.png"), 
caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
label = 'es-depl')
```

\clearpage

## Recruitment{-}

Recruitment is informed by the data from 1980 to 2017 (Table \ref{tab:recrES} and Figure \ref{fig:es-recruits}).  The highest recruitment years occurred in 1999, 2000, 2008, 2013, and 2016. The large 2008 and 2016 year classes, as well as several above average year classes in the early 2010s, contributed to the recent increase in `r spp` biomass. Recruitment is informed by composition data and six relative abundance indices.  The 2015 stock assessment did not estimate deviations from the stock-recruitment curve. While the `r spp` stock has been reduced to levels that theoretically would provide some information on how recruitment compensation changes across spawning biomass levels (i.e., inform the steepness parameter), the assessment model could not adequately estimate a reasonable steepness parameter.  Thus, recruitment is based on a fixed assumption about steepness ($h$ = 0.72) and recruitment variability ($\sigma_R$ = 0.6).

\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
caption = "Estimated time series of age-0 recruits (1000s) for the base model with 95 percent intervals",
label = 'es-recruits')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "recdevs2_withbars.png"), 
caption = "Estimated time series of recruitment deviations",
label = 'es-recdev')
```

## Exploitation status{-}

Fishing intensity, as measured by the SPR rate (1 - SPR), has fluctuated around the estimated SPR rate fishing intensity target of 0.50 (1 - $\text{SPR}_{50\%}$) since 1980. Fishing was at or above the target rate from 1989 to 2005 and has been slightly below it over the past 5 years (Table \ref{tab:exploitES} and Figures \ref{fig:es-1-spr} and \ref{fig:es-phase}). The steepness value of 0.72 indicates that a lower value of SPR (or equivalently a higher fishing intensity than $\text{SPR}_{50\%}$) would be consistent with the biomass-based target of ($\text{SO}_{40\%}$) for sustainable removals. Trends in fishing intensity largely mirrored that of landings until the 1990s, after which recruitment pulses countered the catches somewhat to lower overall fishing intensity (Figure \ref{fig:es-1-spr}). The maximum fishing intensity was `r spr_max` in `r spr_max_yr`, which is well above the target SPR-based harvest rate of 0.50. The current level of `r spr_recent_10[10]` for 2022 is below that target. Fishing intensity over the past decade has ranged between `r min(spr_recent_10)` and `r max(spr_recent_10)` and the exploitation rate (range of `r min(exploit_recent_10)` - `r max(exploit_recent_10)`, Table \ref{tab:exploitES}) has come down since the time series high of 0.12 in 1992. Current estimates indicate that `r spp` spawning output is greater than than the target biomass level ($\text{SO}_{40\%}$), though fishing intensity remains near the target $F_{MSY}$ proxy harvest rate of 1 - $\text{SPR}_{50\%}$ (Figure \ref{fig:es-phase}).

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year for the base model. The management target is plotted as a red horizontal line and values above this reflect harvest in excess of the proxy harvest rate",
label = 'es-1-spr')
```

\clearpage


## Ecosystem considerations{-}

This stock assessment does not explicitly incorporate trophic interactions, habitat factors or environmental factors into the assessment model. More predation, diet and habitat work, and mechanistic linkages to environmental conditions would be needed to incorporate these elements into the stock assessment and should remain a priority. McClure et al. [-@mcclure_vulnerability_2023] report the climate vulnerability for several west coast groundfishes, including `r spp`. `r spp` demonstrated both high biological sensitivity and high climate exposure risk, to give it an overall high vulnerability score to climate change. This result should also be considered with the fact that like many rockfishes, periods of low productivity is not unusually to `r spp` and their extended longevity has historically allowed them to wait for advantageous productivity periods. Additional stressors such as fishing and climate change that possibly truncate longevity could bring significant challenges to population sustainability.

## Reference points{-}

Reference points were based on the rockfish FMSY proxy ($\text{SPR}_{50\%}$), target relative biomass (40%), and estimated selectivity and catch for each fleet (Table \ref{tab:referenceES}). The `r spp` population in Oregon at the start of 2023 is estimated to be 1.38 times (above) the target biomass, and fishing intensity during 2022 is estiamted to be 0.96 times (below) the fishing intensity target (Figure \ref{fig:es-phase}).  The yield values are lower than the previous assessment for similar reference points due to updated life history estimates. The proxy MSY values of management quantities are the most conservative compared to the estimated MSY and MSY relative to 40% biomass. Sustainable total yield, removals, using the proxy $\text{SPR}_{50\%}$ is `r format(model$derived_quants[model$derived_quants$Label == 'Dead_Catch_SPR', 'Value'], digits = 3, scientific = FALSE, big.mark = ',')` mt. The spawning output equivalent to 40 percent of the unfished spawning output ($\text{SO}_{40\%}$) calculated using the SPR target ($\text{SPR}_{50\%}$) was `r format(model$derived_quants[model$derived_quants$Label == 'SSB_SPR', 'Value'], digits = 4, scientific = FALSE, big.mark = ',')` meggs. 

Recent removals have been close to the point estimate of potential long-term yields calculated using an $\text{SPR}_{50\%}$ reference point, though the population size has continued to increase over recent years due to several above average recruitments. The equilibrium estimates of yield relative to biomass based on a steepness value fixed at `r round(model$parameters[model$parameters$Label == "SR_BH_steep","Value"], 2)` are provided in Figure \ref{fig:es-yield}, where vertical dashed lines indicate the estimate of fraction unfished at the start of `r endyr +1` (current) and the estimated management targets calculated based on the relative target biomass (B target), the SPR target, and the maximum sustainable yield (MSY). 

The 2023 spawning biomass relative to unfished equilibrium spawning biomass, based on the 2022 fishing year, is above (55 percent) the management target of 40 percent of unfished spawning output. The relative biomass and the ratio of the estimated SPR to the management target ($\text{SPR}_{50\%}$) across all model years are shown in Figure \ref{fig:es-phase} where warmer colors (red) represent early years and colder colors (blue) represent recent years.  There have been periods where the stock status has decreased below the target and fishing intensity has been higher than the target fishing intensity based on $\text{SPR}_{50\%}$.

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the base model",
label = 'es-phase')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the base case model. Values are based on (the time invariant)
fishery selectivities and with steepness fixed at 0.72",
label = 'es-yield')
```

\input{tex_tables/e_ReferencePoints_ES.tex}

\clearpage

## Management performance{-}

`r spp` removals have generally been below the equivalent ABC-ACL over the recent decade, with the exception of 2017 and 2022 when removals were slightly higher (Table \ref{tab:manageES}). Exploitation on `r spp` increased starting around 1940 and reached a high in the late 1970s. Since that time, catch has mostly fluctuated between 300 and 500 mt per year, with some years exceeding 600 mt. Removals have averaged 483 mt over the past decade. The last ten years of `r spp` acceptable biological catch (ABC) and annual catch limit (ACL) (which are equivalent) has been set, by definition, below the overfishing limit (OFL) (Table \ref{tab:manageES}). Prior to 2017, management specifications were set for `r spp` in California and Oregon waters combined.  The `r spp` OFL has not been exceeded by Oregon removals over the past decade.

## Unresolved problems and major uncertainties{-}

The biggest uncertainty and unresolved conflict is trying to reconcile the signal in the biological data (which suggests a lower population size and status) versus the acoustic and tag surveys (which suggest high stock sizes and status). This is the major issue the current assessment is confronting. Another acoustic-visual survey data point could help resolve how much uncertainty there is in the estimate. The lack of contrast in the biological data, despite large sample sizes, is another barrier to interpreting the current conditions, though given models using only biological data, the signal seems clear that the population could be at a lower stock status.

## Scientific uncertainty{-}

The model-estimated uncertainty around the `r endyr+1` spawning biomass was $\sigma$ = `r round(model$Pstar_sigma,2)` and the uncertainty around the OFL was $\sigma$ = `r round(model$OFL_sigm, 2)`. This is clearly an underestimate of overall uncertainty because of the necessity to fix some life history parameters such as natural mortality and steepness, as well as a lack of explicit incorporation of model structural uncertainty. The alternative states of nature used to bracket uncertainty in the decision table assist with encapsulating model structure uncertainty.

## Harvest Projections and Decision Table{-}

The following text will be modified, as appropriate, after the STAR panel and SSC meeting.

The `r spp` assessment is being considered as a category 2 assessment with a $P^*$ = 0.45, sigma = 0.72, and a time-varying buffer applied to the OFL.  These multipliers are also combined with the rockfish MSY proxy of FSPR=50% MSY and the 40-10 harvest control rule to calculate OFLs, ABCs and ACLs. A twelve year (2023-2034) projection of the reference model using these specifications along with input removals for 2023 and 2024 provided by the Groundfish Management Team is provided in Table \ref{tab:project_ES}. 

Uncertainty in management quantities for the reference model was characterized by exploring various model specifications in a decision table. Initial explorations considered alternative specifications of natural mortality and population scale. Discussion with the STAR panel resulted in selecting a high state of nature based on a model that did not estimate recruitment deviations (i.e., recruitment was based solely on the stock-recruitment relationship). The low state of nature was based on a model that sought to estimate catchability for the acoustic-visual survey. High and low catch streams (rows) were determined by the forecasts, as described above, for each state of nature. Thus the low catch stream is based on the forecast from the low state of nature. The resultant decision table is provided in Table \ref{tab:es-dec-tab}.

 \input{tex_tables/project_ES.tex}

Further details about selecting the decision table states of nature will be added here after the STAR panel.    

Further details summarizing the results presented in the decision table will be added here after the STAR panel.   

\clearpage

 \input{tex_tables/decision_table_es.tex}

\clearpage


## Research and data needs{-}

Recommended avenues for research to help improve future `r spp` stock assessments:

1. The availability of a comprehensive nearshore research survey for `r spp` (as well as for other nearshore rockfish species) was instrumental in providing the stock assessment with an estimate (and associated uncertainty) of population scale.  Given this importance, continue the acoustic-visual fisheries independent coastwide survey to develop a time series. Evaluate survey improvements and analytical refinements as suggested in the SSC methodology review.  Examine alternative spatial modeling approaches to better leverage the available information content toward reducing uncertainty in the absolute estimates of population size. 
2. Improve understanding of broader ecosystem considerations within the context of `r spp` (and other nearshore species) management. Evaluate and develop linkages between `r spp` population dynamics and environmental, oceanographic, and climate variables.  In particular, develop multi-scale models (e.g., species distribution models) that can evaluate spatial patterns (e.g., multi-use areas or closures to fishing) and climate impacts (e.g., growth or distribution shifts) for vulnerable nearshore species. Utilize the growing body of ecosystem information available for the California Current Large Marine Ecosystem, as exemplified in the PFMC IEA report. 
3. Continue work on the investigation into the movement, behavior or mortality of older (> age 10) females to further reconcile their absence in fisheries data. In particular, conduct genetics studies on fish observed off of the continental shelf (middle of the gyre and at sea mounts) to determine their association with the nearshore stocks. 
4. Continue to build evidence for appropriate natural mortality values for females and males. This will help resolve the extent to which dome-shaped age-based selectivity may be occurring for each.
5. Design and conduct research studies to better understand the tradeoffs revealed in this assessment between `r spp` biology and population scale that seem be at odds. If discrepancies can not be uncovered, evaluate management procedures that are as robust as can be to this tradeoff.
6. Conduct early life history studies that provide a better understanding of the ecology and habitats of `r spp` from settlement to age-1.

