\clearpage

# Appendix A: Detailed Fit to Length Composition Data {#app-a}

```{r, results='asis'}
plot_info = read.csv(file.path(mod_loc, "plots", "plotinfotable_for_doc.csv"))
len_fit = plot_info[which(plot_info$category == "LenComp"), ]
plot.vec = grep("comp_lenfit_flt", len_fit$label)
filein = file.path(mod_loc, "plots")

for (a in plot.vec){
	cap = len_fit$caption[a]
	lab = len_fit$label[a]
	add_figure(filein = file.path(filein, len_fit$file[a]), 
			   caption = cap, 
			   label = lab)
}

```

\clearpage

# Appendix B: Fit to Conditional-Age-at-Length Composition Data {#app-b}

```{r, results='asis'}
plot_info = read.csv(file.path(mod_loc, "plots", "plotinfotable_for_doc.csv"))
len_fit = plot_info[which(plot_info$category == "A@LComp"), ]
plot.vec = c(grep("comp_condAALfit_residsflt1mkt0", len_fit$label),
			 grep("comp_condAALfit_residsflt2mkt0", len_fit$label))
find = sort(len_fit$label[plot.vec], index.return = TRUE)$ix
plot.vec = plot.vec[find[2:length(find)]]
filein = file.path(mod_loc, "plots")

for (a in plot.vec){
	cap = len_fit$caption[a]
	lab = len_fit$label[a]
	add_figure(filein = file.path(filein, len_fit$file[a]), 
			   caption = cap, 
			   label = lab)
}

```

\clearpage

# Appendix C: Fit to Conditional-Age-at-Length Composition Data {#app-c}

```{r, results='asis'}
mod_loc = file.path(
   "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/models",
   base_name)
plot_info = read.csv(file.path(mod_loc, "plots", "plotinfotable_for_doc.csv"))
age_fit = plot_info[which(plot_info$category == "A@LComp"), ]
plot.vec = c(grep("comp_condAALfit_Andre_plotsflt1mkt0", age_fit$label),
			 grep("comp_condAALfit_Andre_plotsflt2mkt0", age_fit$label))
find = sort(age_fit$label[plot.vec], index.return = TRUE)$ix
plot.vec = plot.vec[find[2:length(find)]]
filein = file.path(mod_loc, "plots")
cap.in=c("Commerical conditional AAL plot (plot 1 of 8) showing mean age (left panel) and standard deviation (right panel. Shaded areas are 90 percent CIs)",
"Commerical conditional AAL plot (plot 2 of 8)",
"Commerical conditional AAL plot (plot 3 of 8)",
"Commerical conditional AAL plot (plot 4 of 8)",
"Commerical conditional AAL plot (plot 5 of 8)",
"Commerical conditional AAL plot (plot 6 of 8)",
"Commerical conditional AAL plot (plot 7 of 8)",
"Commerical conditional AAL plot (plot 8 of 8)",
"Recreational conditional AAL plot (plot 1 of 4) showing mean age (left panel) and standard deviation (right panel. Shaded areas are 90 percent CIs)",
"Recreational conditional AAL plot (plot 2 of 4)",
"Recreational conditional AAL plot (plot 3 of 4)",
"Recreational conditional AAL plot (plot 4 of 4)"
	)

for (a in plot.vec){
#	cap = age_fit$caption[a]
	cap = cap.in[match(a,plot.vec)]
	lab = age_fit$label[a]
	add_figure(filein = file.path(filein, age_fit$file[a]), 
			   caption = cap, 
			   label = lab)

#base_name = "Reference model"
#mod_loc = file.path(
#   "C:/Users/Jason.Cope/Documents/Github/Vermilion rockfish OR WA assessment 2021/OR/write_up/models",
#   base_name)
#plot_info = read.csv(file.path(mod_loc, "plots/plotinfotable_for_doc.csv"))
#len_fit = plot_info[which(plot_info$category == "A@LComp"), ]
#plot.vec = c(grep("comp_condAALfit_Andre_plotsflt1mkt0", len_fit$label),
#			 grep("comp_condAALfit_Andre_plotsflt2mkt0", len_fit$label))
#find = sort(len_fit$label[plot.vec], index.return = TRUE)$ix
#plot.vec = plot.vec[find[2:length(find)]]
#filein = file.path(mod_loc, "plots")

#for (a in plot.vec){
#	cap = len_fit$caption[a]
#	lab = len_fit$label[a]
#	add_figure(filein = file.path(filein, len_fit$file[a]), 
#			   caption = cap, 
#			   label = lab)
}

```

\clearpage

# Appendix D: Numbers at Age Plot {#app-d}

## Females
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "numbers1_sex1_beg.png"), 
caption = "Female vermilion rockfish mean age over time",
label = 'num_age_females')
```


## Males
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "numbers1_sex2_beg.png"), 
caption = "Male vermilion rockfish mean age over time",
label = 'num_age_males')
```

\clearpage

# Appendix E: Numbers at Length Plot {#app-e}

## Females
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "numbers6_len_sex1.png"), 
caption = "Female vermilion rockfish mean length (cm) over time",
label = 'num_lts_females')
```

\clearpage


## Males
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc, "plots", "numbers6_len_sex2.png"), 
caption = "Male vermilion rockfish mean length over time",
label = 'num_lts_males')
```


\clearpage
