

# Biology

## Maturity and Fecundity

Maturity will borrow values from the Oregon maturity analysis.

Fecundity-at-length is based on Dick et al. (2017) and is similar to what was assumed in the last assessment.

## Length-Weight

The length-weight are given in the figure below. 
Males: W = 0.0000303L^2.843
Females: W = 0.00002881L^2.857

These values are updated with additional samples, yet still similar to those used in the 2015 assessment.

```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "BLK_WA_LW.png"), 
  caption = "Length-weight relationship for black rockfish based on fish sampled in Washington ports.",
  label = 'BLK_WA_LW')

``` 
\

## Age and Growth

Ages are expected to be completed in the next month and will be added to the previous ages in the 2015 assessment. These data will be used to estimate growth outside and inside (as conditional age at length data) the model with the hope we can use internally estimated growth estimates.

The current available samples of age and lengths by sex are shown below. The female to male sex ratio after age 25 is apparent.


```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "WA_FM_growth_curves.png"), 
  caption = "Female and Male age and length data for black rockfish based on fish sampled in Washington ports.",
  label = 'VBGF_FM')

``` 
\


## Natural Mortality

Natural mortality is not directly measured, but has been estimated in past assessments. We will again attempt to estimate natural mortality for both males and females separately. Females demonstrate a lower frequency of occurrence than expected once they get past age 20-25. This issue is seen from California to Alaska, and has been extensively explored in the "kill 'em or hide 'em" hypotheses. These suggest either females are dying off quicker once they reach maturity (i.e., use an elevated natural mortality to kill them in the model) or have age-specific fishing avoidance behavior causing age-based dome-shaped gear selectivity (i.e., hide them from the fishery and create cryptic biomass). These hypotheses were extensively explored in the last assessment. As in the last assessment, we will likely estimate a constant natural mortality for females that is expected to be higher than for males. We will use the [Natural Mortality Tool](https://connect.fisheries.noaa.gov/natural-mortality-tool/) to develop a prior for black rockfish. 
