

# Composition Data

## Commercial length compositions
JASON WILL DO THESE COMPS

## Recreational length compositions
The recreational fishery provides a long time series of length composition by sex. This data is useful for determining fishery selectivity, identifying possible recruitment events and indicating the overall stock status of black rockfish in Washington. Overall, females are slightly larger than males.  

```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "WA-compare-lengths-time-series.png"), 
  caption = "Female and Male length composition data by year for black rockfish based on fish sampled in the Washington sport fishery.",
  label = 'FM_lengths_yrs')

``` 

```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "WA-all-years-density-with-unknown.png"), 
  caption = "Aggregate length composition data by sex for black rockfish based on fish sampled in the Washington sport fishery.",
  label = 'VBGF_FM')

``` 

```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "Sexed_ratios.png"), 
  caption = "Percentage of sexed and unsexed samples in the length composition data for black rockfish based on fish sampled in the Washington sport fishery.",
  label = 'VBGF_FM')

``` 

\


## Survey length compositions

Lisa will add these