

# Biology

## Maturity and Fecundity

Maturity will borrow values from the Oregon maturity analysis.

Fecundity-at-length is based on Dick et al. (2017) and is similar to what was assumed in the last assessment.

## Length-Weight

The length-weight are given in the figure below. 
Males: W = 0.0000303L^2.843
Females: W = 0.00002881L^2.857

These values are updated with additional samples, yet still similar to those used in the 2015 assessment.

```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "BLK_WA_LW.png"), 
  caption = "Length-weight relationship for black rockfish based on fish sampled in Washington ports.",
  label = 'BLK_WA_LW')

``` 
\

## Age and Growth

Ages are expected to be completed in the next month and will be added to the previous ages in the 2015 assessment. These will be used to estimate growth outside and inside the model.

The current available samples of age and lengths by sex are shown below. The female to male sex ratio after age 25 is apparent.


```{r, results = 'asis'}
add_figure(
  filein = file.path(dir, "plots", "WA_FM_growth_curves.png"), 
  caption = "Female and Male age and length data for black rockfish based on fish sampled in Washington ports.",
  label = 'VBGF_FM')

``` 
\


## Natural Mortality

